---
title: "Final project notes"
format: html
---

Overall plan: 1. Use SIPP to create predictive model for household assets (or perhaps for simplicity, just impute the likelihood that a household has assets above limit).

2.  Apply model to ACS to impute household assets. (Do we want to just use gross income, or also net income which excludes deductions? That would be more accurate, but would require more imputation from SIPP).

3.  Merge SNAP policy database data with eligibility requirements to ACS. I'm thinking that the data would be organized with rows = households, and the income and asset threshold variables would be account for household size/state.

4. Calculate eligibility based on 2019 policy environment

5. Simulate different policy scenarios - changes to income and asset thresholds, potentially work requirements. Concern with work requirements is behavioral response

Consider how to deal with debt - how does SNAP eligibility handle that?

"Imputation is not prediction" - consider how we'll want to sample from it

Don't forget the weights

## Federal eligibility requirements
(https://www.congress.gov/crs-product/R42505)

Household definition: person living alone, or people living together who purchase/prepare food together. SIPP does not ask about who purchases/prepares food together, but we can approximate with household relationship variables.

###Income thresholds

**Gross income \<= 130% FPL**

**Net income \<= 100% FPL** 

**Deductions for net income:** 
Standard deduction: A monthly deduction that varies by household size and is indexed for inflation (see below for details). Every applicant household gets this deduction. 
Earned income deduction: 20% of any earned income, in recognition of taxes and work expenses. 
Child support deduction: Any amounts paid out as legally obligated child support, unless already excluded (at the option of the state) from gross income. 
Dependent care deduction: Out-of-pocket dependent care expenses, when related to work or training. 
Excess shelter deduction: Shelter expenses that exceed 50% of net income after all other deductions, typically expenses that exceed about one-third of gross monthly income (see below for limits). Shelter expenses include utility costs, which states may standardize with a standard utility allowance calculation.
Out-of-pocket medical expense deduction: Any medical costs over $35 per month for elderly and disabled members of SNAP unit

### Asset thresholds

\$3,000, or \$4,500 if household includes elderly or disabled member. Includes cash, bank accounts, stocks/bonds, car

Some assets excluded:
Primary home
Most retirement accounts
Resources of SSI/TANF recipients

Many states waive or increase asset thresholds via BBCE

### Work Requirements
**General work requirements:**
Register for work
Accept a suitable job if offered one
Fulfill any work, job search, or training requirements established by administering state SNAP agencies
Provide the administering public assistance agency with sufficient information to allow a determination with respect to their job availability
Not voluntarily quit a job without good cause or reduce work effort below 30 hours a week.

**ABAWD requirements:**
Able-bodied adults without dependents (ABAWDs) must meet some additional requirements:
Work at least 20 hours per week, 
Participate in an employment and training program for at least 20 hours per week, OR
Participate in a state's workfare program

In each 36-month period, ABAWDs are allowed 3 months where they do not meet these requirements.

Some states have waived ABAWD work requirements for the entire state or for certain geographic areas

## 2019 SIPP variables required:
### Household structure and demographics
ERELRPE (household relationship)
RREL* (also household relationship)
ET2REL* (relationship to Type 2 person - lived in household during reference period, but since moved)
TAGE (person's age in years)
ECITIZEN (US citizenship status)
EDISABL (disability status)
ESEX (person's sex)
TAGE_EHC (monthly age during reference period)
ERACE (person's race)
EORIGIN (person's hispanic/latino origin)
EEDUC (person's education level)
RSSI_MNYN (received SSI benefits this month)

### Income
TPTOTINC (total personal income, including from work, assets/investments, and program participation)
TPEARN (total personal earnings)

### Deductions
TAMOUNTPAID (child support payments)
TPAYWK (typical weekly child care payment in fall of reference year)
TRENTMORT (amount household paid for rent or mortgage in December of reference year)
TUTILS (amount the household paid for utilities in December of reference year)
THVAL_RMU (amount household paid for rent, mortgage, and utilities in December of reference year)
TMDPAY (out-of-pocket medical expenses - not clear about reference time)

### Assets (individual-level values are divided equally if asset is shared)
TVAL_AST (total value of all of individual's assets)
TVAL_BANK (total value of all of individual's assets at financial institution)
TVAL_BOND (total value of individual's other interest-earning assets)
TVAL_BUS (total value of individual's shares in businesses)
TVAL_ESAV (total value of individual's educational savings accounts)
TVAL_HOME (total value of individual's primary residence)
TVAL_RENT (total value of individual's rental properties)
TVAL_RE (total value of individual's other real estate holdings)
TVAL_RET (total value of individual's retirement accounts)
TVAL_STMF (total value of individual's stock and mutual fund holdings)
TVAL_VEH (total value of individual's vehicles)
TVAL_OTH (total value of individual's other assets)
TVEH1VAL (total value of first vehicle, household level)
TVEH2VAL (total value of second vehicle, household level)
TVEH3VAL (total value of third vehicle, household level)

### ABAWD work requirements
TMWKHRS (average hours per week at all jobs during the reference month)
EWELACTV2_1 (attended job readiness training)
EWELACTV2_2 (attended job search program)
EWELACTV2_3 (attended training for job skills)
EWELACTV4 (participated in work experience program in exchange for TANF)

### SNAP receipt, for validation
RFSCOV (SNAP receipt any time in reference year)
RFSYN (SNAP receipt this month)

## Set up

```{r}
library(tidyverse)
library(readxl)
library(data.table)

snapdb <- read_xlsx('/Users/JohnnyWilling/Documents/Grad school/Spring 2026/PPOL 6891 Advanced Data Science/final_project/data/SNAPPolicyDatabase.xlsx',
                    sheet = "SNAP Policy Database") |>
  filter(yearmonth == 201907) |>
  select(starts_with("state"), starts_with("bbce"), starts_with("veh"))

sipp2019 <- fread('/Users/JohnnyWilling/Documents/Grad school/Spring 2026/PPOL 6891 Advanced Data Science/final_project/data/pu2019.csv',
                  sep = "|", select = c(
  
  #Common case identification variables
	'SSUID','PNUM','MONTHCODE','ERESIDENCEID','ERELRPE','SPANEL','SWAVE', "TEHC_ST", "TST_INTV",
	
	#The base weight
	'WPFINWGT',
	
	#Common demographics variables, including age at time of interview (TAGE)
	#	and monthly age during the reference period (TAGE_EHC)
	'ESEX','TAGE','TAGE_EHC','ERACE','EORIGIN','EEDUC',
	
	#Example additional variables for analysis
	'TPTOTINC'))

```

## Microsim logic

```{r}
snap_elig <- function(hh_gross_inc, gross_inc_thresh, hh_net_inc, net_inc_thresh, hh_assets, asset_thresh, ) {
  
}

```
