---
title: "Predicting assets"
format: html
editor: visual
---

# Identify variables present in both SIPP and ACS

## Download ACS

```{r}
library(tidycensus)

census_api_key("2d560565103af2c72e380e7fc9b6d7c345cbf0b4", install = TRUE,
               overwrite = TRUE)
```

```{r}
# Filter to 2019 1-year ACS person-level variables
pums_vars <- pums_variables %>%
  filter(year == 2019, survey == "acs1", level == "person")

print(pums_vars)  

# Also look at housing-level variables
pums_vars_hh <- pums_variables %>%
  filter(year == 2019, survey == "acs1", level == "housing")

print(pums_vars_hh)
```

```{r}
# ACS variables
acs_person_vars <- c(
  "AGEP",    # age (-> age_head, has_elderly, has_children)
  "SEX",     # sex
  "RAC1P",   # race
  "HISP",    # Hispanic origin
  "CIT",     # citizenship status
  "NATIVITY",# nativity (native vs foreign-born)
  "SCHL",    # educational attainment
  "ESR",     # employment status
  "WAGP",    # wages/salary income
  "SEMP",    # self-employment income
  "RETP",    # retirement income
  "SSP",     # Social Security income
  "SSIP",    # SSI income
  "PAP",     # public assistance income
  "INTP",    # interest/dividend income
  "PINCP",   # total person income
  "RELSHIPP" # relationship to householder
)

acs_hh_vars <- c(
  "TEN",     # tenure (own vs rent)
  "HINCP",   # household income
  "NP",      # number of persons in household
  "HHT",     # household type
  "VALP",    # property value
  "RNTP",    # monthly rent
  "SMOCP",   # monthly owner costs
  "FS",      # food stamp / SNAP receipt
  "PUMA"     # PUMA geography
)

```

```{r}
# Pull only a few states
acs_raw <- get_pums(
  variables   = c(acs_person_vars, acs_hh_vars),
  state       = c("LA", "TX", "FL", "AL", "CA"),
  survey      = "acs1",
  year        = 2019,
  rep_weights = "both",
  recode      = TRUE
)

# Save locally so no redownloading necessary
saveRDS(acs_raw, file.path("data", "acs_pums_2019.rds"))   
```

```{r}
# Download from data folder
acs <- readRDS("data/acs_pums_2019.rds")
```

## Download SIPP

```{r}
ds <- c("data/pu2019.csv")

all_cols <- names(fread(ds, sep = "|", nrows = 0))
```

```{r}
# SIPP variables
exact_vars <- c(
  # ID variables
  'SSUID', 'PNUM', 'MONTHCODE', 'ERESIDENCEID', 'SPANEL', 'SWAVE',
  'SHHADID', 'THHLDSTATUS', 'RIN_UNIV',
  
  # Base weight
  'WPFINWGT',

  # State of residence
  'TEHC_ST', 'TST_INTV',
  
  # Household structure and demographics 
  'ERELRPE', # (household relationship)
  'TAGE', # (person's age in years)
  'ECITIZEN', # (US citizenship status)
  'EDISABL', # (disability status)
  'ESEX',  # (person's sex)
  'TAGE_EHC', # (monthly age during reference period)
  'ERACE', #(person's race)
  'EORIGIN', # (person's hispanic/latino origin)
  'EEDUC', # (person's education level)
  'RSSI_MNYN', # (received SSI benefits this month)
  
  # Income
  'TPTOTINC', #(total personal income, including from work, assets/investments, and program participation)
  'TPEARN', #(total personal earnings)
  
  # Deductions
  'TAMOUNTPAID', # (child support payments)
  'TPAYWK', # (typical weekly child care payment in fall of reference year)
  'TRENTMORT', # (amount household paid for rent or mortgage in December of reference year)
  'TUTILS', # (amount the household paid for utilities in December of reference year)
  'THVAL_RMU', # (amount household paid for rent, mortgage, and utilities in December of reference year)
  
  # SNAP receipt
  'RFSCOV', # (SNAP receipt any time in reference year)
  'RFSYN' # (SNAP receipt this month)
  
  # Assets
)

patterns <- c(
  "^RREL", #(also household relationship)
  "^ET2REL", # (relationship to Type 2 person - lived in household during reference period, but since moved)
)

pattern_vars <- unique(unlist(lapply(patterns, function(p) grep(p, all_cols, value = TRUE))))

# Combine, keep only cols that exist, and read
keep_vars <- unique(c(exact_vars, pattern_vars))
keep_vars <- keep_vars[keep_vars %in% all_cols]

library(data.table)
pu <- fread(ds, sep = "|", select = keep_vars)
```

```{r}
names(pu) <- toupper(names(pu))
head(pu, 20)
```

## SIPP to ACS Crosswalk

| Concept      | ACS variable | SIPP variable                    |
|--------------|--------------|----------------------------------|
| Age          | `AGEP`       | `TAGE`                           |
| Sex          | `SEX`        | `ESEX`                           |
| Race         | `RAC1P`      | `ERACE`                          |
| Hispanic     | `HISP`       | `EORIGIN`                        |
| Citizenship  | `CIT`        | `ECITIZEN`                       |
| Nativity     | `NATIVITY`   | `ENATIVITY`                      |
| Education    | `SCHL`       | `EEDUC`                          |
| Tenure       | `TEN`        | `ETENURE`                        |
| HH income    | `HINCP`      | `TFTOTINC` (×12 because monthly) |
| SSI          | `SSIP`       | `TSSI_AMT`                       |
| SS income    | `SSP`        | `TSSXAMT`                        |
| SNAP receipt | `FS`         | `RFSCOV, RFSYN`                  |
| HH size      | `NP`         | derived from `SSUID`             |
| State        | `ST`         | `TEHC_ST`                        |

# Define outcome variable in SIPP

Find the asset rules for Louisiana in July 2019

```{r}
# Read in the SNAP policy data
library(readxl)
library(here)

spd <- read_excel(here("data", "SNAPPolicyDatabase.xlsx"),
                  sheet = "SNAP Policy Database")
```

Examine Louisiana July 2019

```{r}
spd |>
  filter(state_pc == "LA") |>
  filter(yearmonth == 201907)
```

Louisiana in July 2019 does not use BBCE.

Cap = 1 means they have combined application project (cap) for SSI recipients

Fingerprints not required

excludes vehicles from asset test (?)
