---
title: "Final project notes"
format: html
editor: 
  markdown: 
    wrap: 72
---

Overall plan:

1.  Use SIPP to create predictive model for household assets (or perhaps
    for simplicity, just impute the likelihood that a household has
    assets above limit).

2.  Apply model to ACS to impute household assets. (Do we want to just
    use gross income, or also net income which excludes deductions? That
    would be more accurate, but would require more imputation from
    SIPP).

3.  Merge SNAP policy database data with eligibility requirements to
    ACS. I'm thinking that the data would be organized with rows =
    households, and the income and asset threshold variables would be
    account for household size/state.

4.  Calculate eligibility based on 2019 policy environment

5.  Simulate different policy scenarios - changes to income and asset
    thresholds, potentially work requirements. Concern with work
    requirements is behavioral response

Consider how to deal with debt - how does SNAP eligibility handle that?

"Imputation is not prediction" - consider how we'll want to sample from
it

Don't forget the weights

## Federal eligibility requirements

(https://www.congress.gov/crs-product/R42505)

Household definition: person living alone, or people living together who
purchase/prepare food together. SIPP does not ask about who
purchases/prepares food together, but we can approximate with household
relationship variables.

###Income thresholds

**Gross income \<= 130% FPL**

**Net income \<= 100% FPL**

**Deductions for net income:** Standard deduction: A monthly deduction
that varies by household size and is indexed for inflation (see below
for details). Every applicant household gets this deduction. Earned
income deduction: 20% of any earned income, in recognition of taxes and
work expenses. Child support deduction: Any amounts paid out as legally
obligated child support, unless already excluded (at the option of the
state) from gross income. Dependent care deduction: Out-of-pocket
dependent care expenses, when related to work or training. Excess
shelter deduction: Shelter expenses that exceed 50% of net income after
all other deductions, typically expenses that exceed about one-third of
gross monthly income (see below for limits). Shelter expenses include
utility costs, which states may standardize with a standard utility
allowance calculation. Out-of-pocket medical expense deduction: Any
medical costs over \$35 per month for elderly and disabled members of
SNAP unit

### Asset thresholds

\$3,000, or \$4,500 if household includes elderly or disabled member.
Includes cash, bank accounts, stocks/bonds, car

Some assets excluded: Primary home Most retirement accounts Resources of
SSI/TANF recipients

Many states waive or increase asset thresholds via BBCE

### Work Requirements

**General work requirements:** Register for work Accept a suitable job
if offered one Fulfill any work, job search, or training requirements
established by administering state SNAP agencies Provide the
administering public assistance agency with sufficient information to
allow a determination with respect to their job availability Not
voluntarily quit a job without good cause or reduce work effort below 30
hours a week.

**ABAWD requirements:** Able-bodied adults without dependents (ABAWDs)
must meet some additional requirements: Work at least 20 hours per week
Participate in an employment and training program for at least 20 hours
per week Participate in a state's workfare program

In each 36-month period, ABAWDs are allowed 3 months where they do not
meet these requirements.

Some states have waived ABAWD work requirements for the entire state or
for certain geographic areas

## 2019 SIPP variables required:

### Household structure and demographics

ERELRPE (household relationship)

RREL (also household relationship)

ET2REL (relationship to Type 2 person - lived in household during
reference period, but since moved) TAGE (person's age in years)

ECITIZEN (US citizenship status)

EDISABL (disability status)

ESEX (person's sex)

TAGE_EHC (monthly age during reference period) ERACE (person's race)
EORIGIN (person's hispanic/latino origin) EEDUC (person's education
level) RSSI_MNYN (received SSI benefits this month)

### Income

TPTOTINC (total personal income, including from work,
assets/investments, and program participation) TPEARN (total personal
earnings)

### Deductions

TAMOUNTPAID (child support payments) TPAYWK (typical weekly child care
payment in fall of reference year) TRENTMORT (amount household paid for
rent or mortgage in December of reference year) TUTILS (amount the
household paid for utilities in December of reference year) THVAL_RMU
(amount household paid for rent, mortgage, and utilities in December of
reference year)

### Assets

### SNAP receipt, for validation

RFSCOV (SNAP receipt any time in reference year) RFSYN (SNAP receipt
this month)

## Set up

Loading in Policy Database

```{r}
library(tidyverse)
library(readxl)
library(data.table)

snapdb <- read_xlsx("SNAPPolicyDatabase.xlsx", sheet = "SNAP Policy Database") |>
  filter(yearmonth == 201907) |>
  select(starts_with("state"), starts_with("bbce"), starts_with("veh"))

## For Work Requirements
link <- "https://fns-prod.azureedge.us/sites/default/files/media/file/FY19-Quarter4-ABAWD-Waiver-Status.pdf"

# Exempt States
state_exempt <- c("Alaska", "District of Columbia", "Louisiana", "New Mexico")

# State fully not exempt
state_nonexempt <- c("Alabama", "Arkansas", "Delaware", "Florida", "Indiana", "Iowa", "Kansas", "Maine", "Mississippi", "Missouri", "Nebraska", "North Carolina", "Oklahoma", "South Carolina", "Texas", "Wisconsin", "Wyoming")

# Modify Rules to Accuont for Work Requirements
snapdb <- snapdb |>
  mutate(work_req = case_when(
    statename %in% state_exempt    ~ "full_exemption",
    statename %in% state_nonexempt ~ "no_exemption",
    TRUE                       ~ "partial_exemption"
  ))

```

Loading in SIPP

```{r}
require("data.table")
require("bit64")
require("dplyr")

ds <- "data/pu2019.csv"

# Read all column names first so we can pattern-match
all_cols <- names(fread(ds, sep = "|", nrows = 0))

exact_vars <- c(
  # ID variables
  'SSUID', 'PNUM', 'MONTHCODE', 'ERESIDENCEID', 'ERELRPE', 'SPANEL', 'SWAVE',
  'SHHADID', 'THHLDSTATUS', 'RIN_UNIV',

  # Base weight
  'WPFINWGT',

  # State of residence
  'TEHC_ST', 'TST_INTV',

  # Demographics
  'ESEX', 'TAGE', 'TAGE_EHC', 'ERACE', 'EORIGIN', 'EEDUC',
  'ECITIZEN', 'ENATIVITY', 'EIMM_STAT',
  'EDISABL', 'RDIS', 'RDIS_ALT', 

  # Household structure
  'TLIVQTR', 'ETENURE', 'EMS',

  # Income
  'TPTOTINC', 'TPEARN', 'TFTOTINC',
  'TUNEMP_AMT', 'TSSI_AMT', 'TSSRX', 'TSSXAMT', 'TTANF_AMT',

  # Deductions
  'TAMOUNTPAID', 'TPAYWK', 'TRENTMORT', 'TUTILS', 'THVAL_RMU',
  'TCHCARE_AMT', 'TMOOP',

  # Assets
  'THNETWORTH', 'ERET_LUMPSUM', 'TDRAW_AMT',

  # Program participation (for eligibility & validation)
  'RFSCOV', 'RSNAP_MNYN', 'RTANF_MNYN', 'RSSI_MNYN', 'RWICYN'
)

patterns <- c(
  "^RREL",   # household relationship
  "^ET2REL", # relationship to Type 2 person
  "^SNAP",
  "^TEHC",
  "^EOWN", "^EJSOWN", "^EJOOWN", "^EOOWN", # asset ownership
  "^TJS", "^TJO", "^VAL",
  "^EVEH", "^TVEH"  # vehicles
)

pattern_vars <- unique(unlist(lapply(patterns, function(p) grep(p, all_cols, value = TRUE))))

keep_vars <- unique(c(exact_vars, pattern_vars))
keep_vars <- keep_vars[keep_vars %in% all_cols]

pu <- fread(ds, sep = "|", select = keep_vars)
names(pu) <- toupper(names(pu))

head(pu, 20)
mean(pu[["TPTOTINC"]], na.rm = TRUE)

# Replicate weights
dw <- "data/rw2019.csv"
rw <- fread(dw, sep = "|")
names(rw) <- toupper(names(rw))

sipp2019 <- left_join(pu, rw, by = c("SSUID", "PNUM", "MONTHCODE"))

head(sipp2019, 20)
```

## Microsim logic

**Filtering SIPP to only include July 2019**

-   Doesn't matter too much when they started (2018, 2019) as long as
    they were in the panel in that month. RIN_UNIV captures this.

```{r}
sipp_0719 <- sipp2019 |>
  filter(RIN_UNIV == 1 & MONTHCODE == 7) #
```

**FPL and Standard Deduction Amounts**

```{r}
# Monthly FPL thresholds for contiguous us and dc (2019)
"https://aspe.hhs.gov/topics/poverty-economic-mobility/poverty-guidelines/prior-hhs-poverty-guidelines-federal-register-references/2019-poverty-guidelines"
fpl_monthly    <- c(1041, 1410, 1778, 2146, 2515, 2883, 3251, 3620)
fpl_additional <- 369  # after 8 people uniform increase

# Standard deduction by household size which only increases up to 6 people, 2019
"https://cvlas.org/wp-content/uploads/2022/08/supp-nutrition-assist-prog-SNAP-2019.pdf"
std_ded_tbl <- c(167, 167, 167, 178, 208, 238)

# Excess shelter deduction cap for non-elderly/non-disabled households, 2019
shelter_cap <- 552

get_fpl <- function(n) {
  fpl_monthly[pmin(n, 8)] + pmax(n - 8, 0) * fpl_additional 
  } # if more than 8 just adds uniform additional ammount

get_std_ded <- function(n) {
  std_ded_tbl[pmin(n, 6)]
  } # more than 6, just applies 6

```

**Aggregate to SNAP filing units**

SNAP filing units are people who live together and purchase/prepare food
together. SIPP does not ask about food purchasing, so we approximate
with household (SSUID).

```{r}
hh_sipp <- sipp_0719 |>
  group_by(SSUID) |>
  summarise(
    # flags
    hh_size = n(),
    has_elderly = any(TAGE >= 60, na.rm = TRUE), 
    has_disabled = any(RDIS == 1, na.rm = TRUE), # simple disability def.
    has_ssi = any(RSSI_MNYN == 1, na.rm = TRUE), # for bbce
    has_tanf = any(RTANF_MNYN == 1, na.rm = TRUE), # Might need to check this
    # Income
    hh_gross_inc = sum(TPTOTINC, na.rm = TRUE), # all forms of earnigns
    hh_earn = sum(TPEARN,   na.rm = TRUE), # earned income
    # Deductions
    child_sup_ded = sum(TAMOUNTPAID, na.rm = TRUE), # how much child support paid
    child_care_ded = sum(TPAYWK * 4, na.rm = TRUE), # child care weekly times 4 for month 
    medical_costs = 0, # struggled to pull variable might need to revisit
    # Shelter costs (in December of reference period)
    shelter_costs = first(na.omit(THVAL_RMU)), # inclues, rent mortgage, utilites for hh
    # Assets 
    net_worth = first(na.omit(THNETWORTH)), # household net worth. 
    state_fips = first(TEHC_ST),
    weight = first(WPFINWGT),
    # Actual SNAP receipiency we can use this to compare, although we are interested in the eligible population. 
    snap_actual = any(RSNAP_MNYN == 1, na.rm = TRUE)
  )
```

### Step 3 — Thresholds, policy merge, net income, and eligibility

```{r}
# Joining in snap eligibility rules to sipp
snap_sipp <- hh_sipp |>
  left_join(snapdb, by = "state_fips") |>
  # Now we have to add in eligibility rules
  mutate(
    # FPL thresholds
    fpl = get_fpl(hh_size),
    gross_limit = fpl * 1.30,
    net_limit   = fpl * 1.00,
    asset_limit = if_else(has_elderly | has_disabled, 4500, 3000),
    # Net income deductions
    std_ded = get_std_ded(hh_size),
    earn_ded  = hh_earn * 0.20,
    net_pre_shelter = pmax(hh_gross_inc - std_ded - earn_ded 
                           - child_sup_ded - child_care_ded, 0),
    shelter_excess = pmax(shelter_costs - 0.50 * net_pre_shelter, 0),
    shelter_ded = if_else(has_elderly | has_disabled, 
                          shelter_excess, pmin(shelter_excess, shelter_cap)),
    medical_ded = if_else(has_elderly | has_disabled, pmax(medical_costs - 35, 0), 0),
    net_inc = pmax(net_pre_shelter - shelter_ded - medical_ded, 0),
    # Categorical eligibility: SSI/TANF recipients skip income and asset tests
    cat_elig = has_ssi | has_tanf,
    # BBCE: states may waive asset test and raise gross income limit
    # Logic of this next section is do they have bbce, if so use that income limit and asset test. 
    bbce_on = !is.na(bbce) & bbce == 1,
    gross_limit_eff = if_else(
      bbce_on & bbce_inclmt > 0,
      fpl * (bbce_inclmt / 100),
      gross_limit),
    asset_test_on   = !(bbce_on & !is.na(bbce_asset) & bbce_asset == 1),
    asset_limit_eff = case_when(
      bbce_on & bbce_a_amt > 0 ~ bbce_a_amt * 1000,
      TRUE ~ asset_limit),
    # Individual tests
    gross_elig = hh_gross_inc <= gross_limit_eff,
    net_elig   = net_inc      <= net_limit,
    asset_elig = !asset_test_on | cat_elig | (net_worth <= asset_limit_eff),
    # Work requirements: single, non-elderly, non-disabled, age 18–49
    # Simplified to household level — ideally flagged at person level
    has_abawd  = hh_size == 1 & !has_elderly & !has_disabled & !has_ssi & !has_tanf,
    work_elig  = work_req %in% c("full_exemption") | !has_abawd,
    # Final eligibility
    snap_eligible = (cat_elig | (gross_elig & net_elig)) & asset_elig & work_elig)

snap_eligible <- snap_sipp |>
  filter(snap_eligible == TRUE)
```

**Basic Descriptive Statistics of how many families**

```{r}
# Unweighted numbers
snap_sipp |>
  summarize(
    n_eligible = sum(snap_eligible, na.rm = TRUE),
    n_total = n(),
    pct_eligible = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1)
  )

# Weighted Numbers
snap_sipp |>
  summarize(
    n_eligible_weighted = sum(snap_eligible * weight, na.rm = TRUE),
    n_total_weighted    = sum(weight, na.rm = TRUE),
    pct_eligible        = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1)
  )
```


BIG NOTES: 
CURRENTLY OVERCOUNTING
NEED TO ADD IN WORK INFORMATION HOURS, ETC, 
NEED TO EDIT IN REVISED ASSET STUFF
MIGHT NEED TO FIX MEDICAL EXPENSES

