---
title: "SNAP Eligibility Microsimulation"
format:
  html:
    embed-resources: true
editor:
  markdown:
    wrap: 72
execute:
  cache: true
  cache.lazy: false
  warning: false
  message: false
---

## Overview

This file is fully self-contained. It loads all data from scratch,
builds household-level SNAP filing units, defines a vectorized
eligibility function, and runs baseline and counterfactual policy
scenarios.

## Libraries

```{r}
#| cache: false
library(tidyverse)
library(readxl)
library(data.table)
library(bit64)
library(skimr)
library(sf)
library(tigris)
library(scales)
library(ggrepel)
```

## Load SNAP Policy Database

[The SNAP Policy
Database](https://www.ers.usda.gov/data-products/snap-policy-data-sets),
created by the Food and Nutrition Service, contains data on state-level
SNAP policies, such as broad-based categorical eligibility (BBCE) rules
related to income and asset tests. We use SNAP Policy Database data from
July 2019 and merge in information on exemptions to work requirement
time limits for able-bodied adults without dependents (ABAWDs), also
[available via
FNS](https://fns-prod.azureedge.us/sites/default/files/media/file/FY19-Quarter3-ABAWD-Waiver-Status.pdf)
for July 2019.

```{r}
snapdb <- read_xlsx("data/SNAPPolicyDatabase.xlsx", sheet = "SNAP Policy Database") |>
  filter(yearmonth == 201907) |>
  select(starts_with("state"), starts_with("bbce"), starts_with("veh"))

# Work requirement tiers based on FY2019 Q4 ABAWD waiver status
state_exempt <- c("Alaska", "District of Columbia", "Louisiana", "New Mexico")

state_nonexempt <- c("Alabama", "Arkansas", "Delaware", "Florida", "Indiana",
                     "Iowa", "Kansas", "Maine", "Mississippi", "Missouri",
                     "Nebraska", "North Carolina", "Oklahoma", "South Carolina",
                     "Texas", "Wisconsin", "Wyoming")

snapdb <- snapdb |>
  mutate(work_req = case_when(
    statename %in% state_exempt ~ "full_exemption",
    statename %in% state_nonexempt ~ "no_exemption",
    TRUE ~ "partial_exemption"))
```

## Load SIPP 2019
The core dataset for our microsimulation comes from the Survey of Income and Program Participation (SIPP), a longitudinal panel survey of US households that contains rich data on income, assets, and participation in public programs. We use data from 2019, which includes the 2018 and 2019 SIPP panels, available from the [Census website](https://www.census.gov/programs-surveys/sipp/data/datasets/2019-data/2019.html).

```{r}
#| cache: false
ds <- "data/pu2019.csv"

all_cols <- names(fread(ds, sep = "|", nrows = 0))

exact_vars <- c(
  # ID variables
  'SSUID', 'PNUM', 'MONTHCODE', 'ERESIDENCEID', 'ERELRPE', 'SPANEL', 'SWAVE',
  'SHHADID', 'THHLDSTATUS', 'RIN_UNIV',

  # Base weight
  'WPFINWGT',

  # State of residence
  'TEHC_ST', 'TST_INTV',

  # Demographics
  'ESEX', 'TAGE', 'TAGE_EHC', 'ERACE', 'EORIGIN', 'EEDUC',
  'ECITIZEN', 'ENATIVITY', 'EIMM_STAT', 'TIMSTAT', 'TYRENTRY',
  'EDISABL', 'RDIS', 'RDIS_ALT',

  # Household structure
  'TLIVQTR', 'ETENURE', 'EMS',

  # Income
  'TPTOTINC', 'TPEARN', 'TFTOTINC',
  'TUNEMP_AMT', 'TSSI_AMT', 'TSSRX', 'TSSXAMT', 'TTANF_AMT',

  # Deductions
  'TAMOUNTPAID', 'TPAYWK', 'TRENTMORT', 'TUTILS', 'THVAL_RMU',
  'TCHCARE_AMT', 'TMDPAY',

  # Assets (individual-level; shared assets divided equally)
  'TVAL_AST', 'TVAL_BANK', 'TVAL_BOND', 'TVAL_BUS', 'TVAL_ESAV',
  'TVAL_HOME', 'TVAL_RENT', 'TVAL_RE', 'TVAL_RET',
  'TVAL_STMF', 'TVAL_VEH', 'TVAL_OTH',

  # Program participation
  'RFSCOV', 'RSNAP_MNYN', 'RTANF_MNYN', 'RSSI_MNYN', 'RWICYN',

  # ABAWD work requirement variables
  'TMWKHRS', 'EWELACTV2_1', 'EWELACTV2_2', 'EWELACTV2_3', 'EWELACTV4'
)

patterns <- c(
  "^RREL", "^ET2REL", "^SNAP", "^TEHC",
  "^EOWN", "^EJSOWN", "^EJOOWN", "^EOOWN",
  "^TJS", "^TJO", "^VAL", "^EVEH", "^TVEH"
)

pattern_vars <- unique(unlist(lapply(patterns, function(p) grep(p, all_cols, value = TRUE))))
keep_vars    <- unique(c(exact_vars, pattern_vars))
keep_vars    <- keep_vars[keep_vars %in% all_cols]

pu <- fread(ds, sep = "|", select = keep_vars)
names(pu) <- toupper(names(pu))

dw <- "data/rw2019.csv"
rw <- fread(dw, sep = "|")
names(rw) <- toupper(names(rw))

sipp2019 <- left_join(pu, rw, by = c("SSUID", "PNUM", "MONTHCODE"))
```

## Filter to July 2019 and generate new SNAP unit ID

Households are uniquely identified in the SIPP by SSUID and SHHADID, but some households may contain multiple SNAP units, defined as people who live and purchase food together. The SIPP does not directly ask about this, but we approximate it based on household relationships. We assume that family members and partners living together are part of the same SNAP unit, while roommates/boarders/other non-family members are not.

```{r}
sipp_0719 <- sipp2019 |>
  filter(RIN_UNIV == 1 & MONTHCODE == 7) |>
  mutate(hhid = paste0(as.character(SSUID), as.character(SHHADID))) |>
  group_by(hhid) |>
  mutate(snap_split = ERELRPE %in% 16:18, # splitting cases that are roommates/boarders/other non-family into separate SNAP unit
         split_id = if_else(snap_split, cumsum(snap_split), NA_integer_),
         snapuid = case_when(!snap_split ~ 1,
         snap_split ~ split_id + 1)
) |> ungroup() |>
  mutate(snap_hhid = paste0(hhid, snapuid))

# Brief checks to make sure split went correctly
sipp_0719 |>
  group_by(hhid) |>
  summarise(n_snap_units = n_distinct(snap_hhid)) |>
  count(n_snap_units)

sipp_0719 |>
  group_by(hhid) |>
  filter(n_distinct(snap_hhid) > 1) |>
  arrange(hhid, snap_hhid) |>
  select(hhid, ERELRPE, snap_hhid) |>
  ungroup() |>
  slice_head(n = 50)

```

## FPL and Deduction Limits

```{r}
# 2019 monthly FPL (contiguous 48 states and DC)
# https://aspe.hhs.gov/topics/poverty-economic-mobility/poverty-guidelines/prior-hhs-poverty-guidelines-federal-register-references/2019-poverty-guidelines 
fpl_monthly    <- c(1041, 1410, 1778, 2146, 2515, 2883, 3251, 3620)
fpl_additional <- 369

# 2019 standard deductions by household size (caps at 6)
# https://cvlas.org/wp-content/uploads/2022/08/supp-nutrition-assist-prog-SNAP-2019.pdf
std_ded_tbl <- c(167, 167, 167, 178, 208, 238)

# Excess shelter deduction cap for non-elderly/disabled households
shelter_cap <- 552

get_fpl <- function(n){
  fpl_monthly[pmin(n, 8)] + pmax(n - 8, 0) * fpl_additional
}
get_std_ded <- function(n){
  std_ded_tbl[pmin(n, 6)]
}
```

## Aggregate to Household-Level SNAP Filing Units

SNAP filing units are people who live and purchase food together. Given
that SIPP doesn't account for people who consume food together, we use
SSUID to account for the pseudo-economic unit.

[It's important to note the SIPP is very well known for undercounting
income
information.](https://www.aeaweb.org/articles?id=10.1257/jep.29.4.199)

Since we are constructing our own SNAP households based on SNAP
methodology (based on those who share food), we use the SIPP User
Guide's recommendation to create household weights by simply taking a
reference person's weight to represent the snap household. More
information on this methodology can be found in [Section 7.3.3 of the
user
guide.](https://www2.census.gov/programs-surveys/sipp/tech-documentation/methodology/2024_SIPP_Users_Guide.pdf)

Citation: U.S. Census Bureau. (2025). *2024 Survey of Income and Program
Participation Users' Guide* (July 2025). Section 7.3.3, p. 161.

```{r}
# Aggregation
hh_sipp <- sipp_0719 |>
  group_by(snap_hhid) |>
  summarize(
    hh_size = n(),
    has_elderly = any(TAGE >= 60, na.rm = TRUE),
    has_disabled = any(RDIS == 1, na.rm = TRUE),
    has_ssi = all(RSSI_MNYN == 1, na.rm = TRUE),
    has_tanf = all(RTANF_MNYN == 1, na.rm = TRUE),
    # Income
    hh_gross_inc = sum(TPTOTINC, na.rm = TRUE) * 1.25, # sipp undercounts apply this factor
    hh_earn = sum(TPEARN, na.rm = TRUE) * 1.25, # sipp undercounts apply this factor
    # Deductions
    child_sup_ded = sum(TAMOUNTPAID, na.rm = TRUE),
    child_care_ded = sum(TPAYWK, na.rm = TRUE) * 4,    
    medical_costs = sum(TMDPAY, na.rm = TRUE),
    # Shelter (December of reference year as proxy)
    shelter_costs = coalesce(first(na.omit(THVAL_RMU)), 0),
    # Countable assets
    countable_assets = sum(TVAL_BANK, na.rm = TRUE) + sum(TVAL_BOND, na.rm = TRUE) +
                       sum(TVAL_BUS,  na.rm = TRUE) + sum(TVAL_STMF, na.rm = TRUE) +
                       sum(TVAL_RENT, na.rm = TRUE) + sum(TVAL_RE,   na.rm = TRUE) +
                       sum(TVAL_OTH,  na.rm = TRUE),
    vehicle_assets = sum(TVAL_VEH, na.rm = TRUE),
    tveh1_val = first(TVEH1VAL),
    tveh2_val = first(TVEH2VAL),
    # ABAWD flags
    has_children = any(TAGE < 18, na.rm = TRUE),      
    youngest_child_age = if (any(TAGE < 26, na.rm = TRUE)) 
                           as.numeric(min(TAGE[TAGE < 26], na.rm = TRUE))
                         else NA_real_,
    has_abawd_age = any(TAGE >= 18 & TAGE <= 49 & # baseline 18-49 flag
                       (is.na(RDIS) | RDIS != 1), na.rm = TRUE),
    has_adult_nondis_50_59 = any(TAGE >= 50 & TAGE <= 59 &
                                (is.na(RDIS) | RDIS != 1), na.rm = TRUE),
    has_adult_nondis_60_65 = any(TAGE >= 60 & TAGE <= 65 &
                                (is.na(RDIS) | RDIS != 1), na.rm = TRUE),
    max_work_hrs = max(c(0, TMWKHRS[TAGE >= 18 &         
                       (is.na(RDIS) | RDIS != 1)]), na.rm = TRUE),
    has_cit_or_elig_noncit = any(ECITIZEN == 1 | (TIMSTAT == 1 & TYRENTRY <= 2014), na.rm = TRUE), 
    state_fips = first(TEHC_ST),
    weight = first(WPFINWGT[!is.na(ERELRPE) & ERELRPE == min(ERELRPE, na.rm = TRUE)]),
    snap_actual = any(RSNAP_MNYN == 1, na.rm = TRUE)
  )
```

## Eligibility Function

State-level BBCE and vehicle rules are looked up internally from
`snapdb` using `state_fips`. We hardcoded three national or other
baseline level policy levers which can then be modified for simulation.

They are income, asset, and abawd work policy levers.

```{r}
snap_elig_sim <- function(df,
    # Note all of these can be changed as simulations: Income 
    gross_pct = 1.30,     # gross income limit as fraction of FPL (federal default: 1.30)
    net_pct = 1.00,     # net income limit as fraction of FPL   (federal default: 1.00)
    use_bbce= TRUE,     # apply state BBCE income extensions
    # Asset 
    elim_asset = FALSE,  # TRUE = no asset test for anyone
    asset_limit = NULL,   # set a uniform asset limit in dollars; NULL = federal/BBCE default
    # Work 
    work_req = NULL, # NULL = use state rules; "full_exemption"; "no_exemption"
    abawd_max_age = 49,   # upper age of ABAWD range (federal default: 49)
    min_work_hrs = 20,   # minimum weekly hours to meet work requirement 
    dep_age = 18    # child must be under this age to exempt adult from ABAWD 
) { with(df, 
         {
  # State snap policy
  sp <- snapdb[match(state_fips, snapdb$state_fips), ]

  # FPL thresholds
  fpl       <- get_fpl(hh_size)
  gross_lmt <- fpl * gross_pct
  net_lmt   <- fpl * net_pct
  fed_asset_lmt <- if_else(has_elderly | has_disabled, 3500, 2250)
  if (!is.null(asset_limit)) fed_asset_lmt <- rep(asset_limit, length(hh_size))

  # Net income calculation
  std_ded <- get_std_ded(hh_size)
  earn_ded <- hh_earn * 0.20
  net_pre_shelter <- pmax(hh_gross_inc - std_ded - earn_ded - child_sup_ded - child_care_ded, 0)
  shelter_excess  <- pmax(shelter_costs - 0.50 * net_pre_shelter, 0)
  shelter_ded <- if_else(has_elderly | has_disabled,
                             shelter_excess, pmin(shelter_excess, shelter_cap))
  medical_ded <- if_else(has_elderly | has_disabled, pmax(medical_costs - 35, 0), 0)
  net_inc <- pmax(net_pre_shelter - shelter_ded - medical_ded, 0)

  # Categorical eligibility (SSI/TANF bypasses income and asset tests)
  cat_elig <- has_ssi | has_tanf

  # BBCE — look up from state policy; disabled when use_bbce = FALSE
  bbce_on <- use_bbce & !is.na(sp$bbce) & sp$bbce == 1
  bbce_applies <- bbce_on & !(coalesce(sp$bbce_child, 0L) == 1L & !has_children)

  gross_lmt_bbce <- if_else(bbce_applies & coalesce(sp$bbce_inclmt, -9L) > 0,
                             fpl * (sp$bbce_inclmt / 100), gross_lmt)
  gross_lmt_eld <- if_else(bbce_applies & coalesce(sp$bbce_elddisinclmt, -9L) > 0,
                             fpl * (sp$bbce_elddisinclmt / 100), gross_lmt_bbce)
  gross_lmt_eff <- if_else(has_elderly | has_disabled, gross_lmt_eld, gross_lmt_bbce)

  asset_test_on <- !(bbce_applies & !is.na(sp$bbce_asset) & sp$bbce_asset == 1)
  asset_lmt_eff <- case_when(
    bbce_applies & coalesce(sp$bbce_a_amt, 0) > 0 ~ sp$bbce_a_amt * 1000,
    TRUE ~ fed_asset_lmt
  )
  if (elim_asset) asset_test_on <- rep(FALSE, length(hh_size))

  # Vehicle countable value — look up state vehicle exclusion rules
  vehicle_countable <- case_when(
    sp$vehexclall == 1 ~ 0,
    sp$vehexclone == 1 ~ pmax(vehicle_assets - coalesce(tveh1_val, 0), 0),
    TRUE ~ vehicle_assets
  )
  net_worth <- countable_assets + vehicle_countable

  # Eligibility tests
  gross_elig <- (has_elderly | has_disabled) | hh_gross_inc <= gross_lmt_eff
  net_elig   <- net_inc      <= net_lmt
  asset_elig <- !asset_test_on | cat_elig | (net_worth <= asset_lmt_eff)

  eff_work_req <- if (!is.null(work_req)) rep(work_req, length(hh_size)) else sp$work_req
  has_abawd_age_eff  <- has_abawd_age |
                        (abawd_max_age >= 55 & has_adult_nondis_50_59) |
                        (abawd_max_age >= 60 & has_adult_nondis_60_65)
  has_qualifying_dep <- !is.na(youngest_child_age) & youngest_child_age < dep_age
  meets_work_eff <- max_work_hrs >= min_work_hrs
  has_abawd <- has_abawd_age_eff & !has_qualifying_dep & !has_ssi & !has_tanf
  work_elig <- eff_work_req == "full_exemption" | !has_abawd | meets_work_eff

  citizen_elig <- coalesce(has_cit_or_elig_noncit, TRUE) | cat_elig

  (cat_elig | (gross_elig & net_elig)) & asset_elig & work_elig & citizen_elig
})}

```

## Scenario Functions

```{r}
run_scenario <- function(data, scenario_name, ...) {
  data |>
    mutate(snap_eligible = snap_elig_sim(pick(everything()), ...),
           scenario = scenario_name)
}

summarize_scenario <- function(data) {
  data |>
    ungroup() |>
    summarize(
      scenario = first(scenario),
      n_eligible_weighted = sum(snap_eligible * weight, na.rm = TRUE),
      n_total_weighted = sum(weight, na.rm = TRUE),
      pct_eligible= round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1))
}
```

## Run Scenarios at National Level

Roughly [82% of SNAP eligible
households](https://www.fns.usda.gov/research/snap/state-participation-rates/2019#:~:text=Key%20Findings,Page%20updated:%20April%2018%2C%202025)
participate in snap in 2019. [July 2019 had 19 million
households](https://www.fns.usda.gov/pd/supplemental-nutrition-assistance-program-snap).
As such assuming the SIPP properly captures income, asset information,
then we should have roughly 23.5 million households.

```{r}
# Baseline: 2019 state rules as-is
baseline <- run_scenario(hh_sipp, "Baseline (2019)")

# Income
income_200pct  <- run_scenario(hh_sipp, "Gross limit 200% FPL",
                               gross_pct = 2.00)

income_no_bbce <- run_scenario(hh_sipp, "Federal rules only (no BBCE)",
                               use_bbce = FALSE)

# Assets
asset_elim <- run_scenario(hh_sipp, "Eliminate asset test",
                           elim_asset = TRUE)

asset_10k  <- run_scenario(hh_sipp, "Raise asset limit to $10k",
                           asset_limit = 10000)

# Work requirement scenarios 
work_waive <- run_scenario(hh_sipp, "Waive ABAWD nationally",
                             work_req = "full_exemption")

work_enforce <- run_scenario(hh_sipp, "Enforce ABAWD nationally",
                             work_req = "no_exemption")

work_30hrs <- run_scenario(hh_sipp, "30-hour requirement", min_work_hrs = 30)

work_dep5    <- run_scenario(hh_sipp, "Dep age cutoff 5", dep_age = 5)

work_age65 <- run_scenario(hh_sipp, "ABAWD age 18-65", abawd_max_age = 65)
```

## National Results

```{r}
results <- bind_rows(
  summarize_scenario(baseline),
  summarize_scenario(income_200pct),
  summarize_scenario(income_no_bbce),
  summarize_scenario(asset_elim),
  summarize_scenario(asset_10k),
  summarize_scenario(work_waive),
  summarize_scenario(work_enforce),
  summarize_scenario(work_30hrs),
  summarize_scenario(work_dep5),
  summarize_scenario(work_age65),
) |>
  mutate(
    change_from_baseline = n_eligible_weighted - n_eligible_weighted[scenario == "Baseline (2019)"],
    pct_change = round(100 * change_from_baseline / n_eligible_weighted[scenario == "Baseline (2019)"], 1)
  )

results
```

## Run Scenarios at the State Level

```{r}
summarize_by_state <- function(data) {
  data |>
    group_by(state_fips, scenario) |>
    summarize(
      n_eligible_weighted = sum(snap_eligible * weight, na.rm = TRUE),
      n_total_weighted    = sum(weight, na.rm = TRUE),
      pct_eligible        = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1),
      .groups = "drop"
    )
}
```

```{r}
# Baseline
baseline_state <- hh_sipp |>
  run_scenario("Baseline") |>
  summarize_by_state()

# Income
twohundred_state  <- hh_sipp |>
  run_scenario("Gross limit 200% FPL", gross_pct = 2.00) |>
  summarize_by_state()

no_bbce_state <- hh_sipp |>
  run_scenario("Federal rules only (no BBCE)", use_bbce = FALSE) |>
  summarize_by_state()

# Assets
asset_elim_state <- hh_sipp |>
  run_scenario("Eliminate asset test", elim_asset = TRUE) |>
  summarize_by_state()

asset_10k_state  <- hh_sipp |>
  run_scenario("Raise asset limit to $10k", asset_limit = 10000) |>
  summarize_by_state()

# Work requirement scenarios 
work_waive_state <- hh_sipp |>
  run_scenario("Waive ABAWD nationally", work_req = "full_exemption") |>
  summarize_by_state()

work_enforce_state <- hh_sipp |>
  run_scenario("Enforce ABAWD nationally", work_req = "no_exemption") |>
  summarize_by_state()

work_30hrs_state <- hh_sipp |>
  run_scenario("30-hour requirement", min_work_hrs = 30) |>
  summarize_by_state()

work_dep5_state <- hh_sipp |>
  run_scenario("Dep age cutoff 5", dep_age = 5) |>
  summarize_by_state()

work_age65_state <- hh_sipp |>
  run_scenario("ABAWD age 18-65", abawd_max_age = 65) |>
  summarize_by_state()
```

## State Level Results

```{r}
state_results <- bind_rows(baseline_state, twohundred_state,
                           no_bbce_state, asset_elim_state, asset_10k_state,
                           work_waive_state, work_enforce_state, work_30hrs_state,
                           work_dep5_state, work_age65_state) |>
  group_by(state_fips) |>
  mutate(change_from_baseline = n_eligible_weighted - n_eligible_weighted[scenario == "Baseline"],
         pct_change = round(100 * change_from_baseline / n_eligible_weighted[scenario == "Baseline"],
                            1)) |>
  ungroup() |>
  arrange(state_fips)

state_results
```

# Visualize State Level Results

```{r}
visualize_scenario <- function(df, scen) {
  # State geometry from tigris
  states <- states(cb = TRUE, resolution = "20m") |>
    st_transform(5070) |>
    shift_geometry()
  
  # Separate out small states that will need custom labels
  small_states <- c("VT","NH","MA","RI","CT","NJ","DE","MD","DC")
  
  # Merge state geometry with eligibility microsim data
  plot_data <- df |> 
    filter(scenario == scen) |>
    mutate(STATEFP = str_pad(state_fips, width = 2, side = "left", pad = "0")) |>
    left_join(states, by = "STATEFP") |>
    st_as_sf() |>
    mutate(is_small = STUSPS %in% small_states)
  
  # Label positions and table for small states
  centroids <- st_centroid(plot_data)
  
  coords <- centroids %>%
    mutate(
      x = st_coordinates(.)[,1],
      y = st_coordinates(.)[,2]
    ) %>%
    st_drop_geometry()
  
  table_positions <- coords |>
    filter(is_small) |>
    arrange(desc(y)) |> 
    mutate(
      table_x = max(coords$x) + 6e5,
      table_y = seq(max(coords$y), length.out = n(), by = -2e5)
    )
  
  coords <- coords |>
    left_join(table_positions |> select(STUSPS, table_x, table_y), by = "STUSPS")
  
  # Make plot
  ggplot(plot_data) +
    geom_sf(aes(fill = pct_eligible), color = "white") +
    # Labels for large states
    geom_text(
      data = coords |> filter(!is_small),
      aes(x = x, y = y, label = pct_eligible),
      size = 3
    ) +
    # Segments for small states
    geom_segment(
      data = coords |> filter(is_small),
      aes(x = x, y = y, xend = table_x, yend = table_y),
      color = "gray50"
    ) +
    # Separate labels for small states
    geom_text(
      data = coords |> filter(is_small),
      aes(x = table_x, y = table_y, 
          label = paste0(STUSPS, ": ", pct_eligible)),
      hjust = 0,
      size = 3
    ) +
    scale_fill_gradient(low = "#BAE1FF", high = "#0B538A") +
    coord_sf(clip = "off") +
    labs(title = scen, fill = "Percent of households SNAP eligible") +
    theme_void() +
    theme(
      plot.title = element_text(hjust = 0.5),
      plot.margin = margin(10, 100, 10, 10),
      legend.position = "bottom"
    )
    
}


```

```{r}
visualize_scenario(state_results, scen = "Baseline")

visualize_scenario(state_results, scen = "Gross limit 200% FPL")

visualize_scenario(state_results, scen = "Federal rules only (no BBCE)")

visualize_scenario(state_results, scen = "Eliminate asset test")

visualize_scenario(state_results, scen = "Raise asset limit to $10k")

visualize_scenario(state_results, scen = "Waive ABAWD nationally")

visualize_scenario(state_results, scen = "Enforce ABAWD nationally")

visualize_scenario(state_results, scen = "30-hour requirement")

visualize_scenario(state_results, scen = "Dep age cutoff 5")

visualize_scenario(state_results, scen = "ABAWD age 18-65")

```
