---
title: "SNAP Eligibility Microsimulation"
format:
  html:
    code-fold: true
    embed-resources: true
editor:
  markdown:
    wrap: 72
execute:
  cache: true
  cache.lazy: false
  warning: false
  message: false
bibliography: references.bib
---

## Overview

This file is fully self-contained. It loads all data from scratch,
builds household-level SNAP filing units, defines a vectorized
eligibility function, and runs baseline and counterfactual policy
scenarios.

## Libraries

```{r}
#| warning: false
#| message: false
#| cache: false
library(tidyverse)
library(readxl)
library(data.table)
library(bit64)
library(tigris)
library(sf)
```

## Load SNAP Policy Database

The SNAP Policy Database, created by the Food and Nutrition Service, 
contains data on state-level SNAP policies, such as broad-based categorical 
eligibility (BBCE) rules related to income and asset tests [@snapdb]. We use SNAP 
Policy Database data from July 2019 and merge in information on exemptions 
to work requirement time limits for able-bodied adults without dependents
(ABAWDs), also available via FNS for July 2019 [@fns2019].

```{r}
snapdb <- read_xlsx("data/SNAPPolicyDatabase.xlsx", sheet = "SNAP Policy Database") |>
  filter(yearmonth == 201907) |>
  select(starts_with("state"), starts_with("bbce"), starts_with("veh"))

# Work requirement tiers based on FY2019 Q4 ABAWD waiver status
state_exempt <- c("Alaska", "District of Columbia", "Louisiana", "New Mexico")

state_nonexempt <- c("Alabama", "Arkansas", "Delaware", "Florida", "Indiana",
                     "Iowa", "Kansas", "Maine", "Mississippi", "Missouri",
                     "Nebraska", "North Carolina", "Oklahoma", "South Carolina",
                     "Texas", "Wisconsin", "Wyoming")

snapdb <- snapdb |>
  mutate(work_req = case_when(
    statename %in% state_exempt ~ "full_exemption",
    statename %in% state_nonexempt ~ "no_exemption",
    TRUE ~ "partial_exemption"))
```

## Load SIPP 2019
The core dataset for our microsimulation comes from the Survey of Income and Program Participation (SIPP), a longitudinal panel survey of US households that contains rich data on income, assets, and participation in public programs. We use data from 2019, which includes the 2018 and 2019 SIPP panels, available from the Census website [@sipp2019].

```{r}
#| cache: false
ds <- "data/pu2019.csv"

all_cols <- names(fread(ds, sep = "|", nrows = 0))

exact_vars <- c(
  # ID variables
  'SSUID', 'PNUM', 'MONTHCODE', 'ERESIDENCEID', 'ERELRPE', 'SPANEL', 'SWAVE',
  'SHHADID', 'THHLDSTATUS', 'RIN_UNIV',

  # Base weight
  'WPFINWGT',

  # State of residence
  'TEHC_ST', 'TST_INTV',

  # Demographics
  'ESEX', 'TAGE', 'TAGE_EHC', 'ERACE', 'EORIGIN', 'EEDUC',
  'ECITIZEN', 'ENATIVITY', 'EIMM_STAT', 'TIMSTAT', 'TYRENTRY',
  'EDISABL', 'RDIS', 'RDIS_ALT',

  # Household structure
  'TLIVQTR', 'ETENURE', 'EMS',

  # Income
  'TPTOTINC', 'TPEARN', 'TFTOTINC',
  'TUNEMP_AMT', 'TSSI_AMT', 'TSSRX', 'TSSXAMT', 'TTANF_AMT',

  # Deductions
  'TAMOUNTPAID', 'TPAYWK', 'TRENTMORT', 'TUTILS', 'THVAL_RMU',
  'TCHCARE_AMT', 'TMDPAY',

  # Assets (individual-level; shared assets divided equally)
  'TVAL_AST', 'TVAL_BANK', 'TVAL_BOND', 'TVAL_BUS', 'TVAL_ESAV',
  'TVAL_HOME', 'TVAL_RENT', 'TVAL_RE', 'TVAL_RET',
  'TVAL_STMF', 'TVAL_VEH', 'TVAL_OTH',

  # Program participation
  'RFSCOV', 'RSNAP_MNYN', 'RTANF_MNYN', 'RSSI_MNYN', 'RWICYN',

  # ABAWD work requirement variables
  'TMWKHRS', 'EWELACTV2_1', 'EWELACTV2_2', 'EWELACTV2_3', 'EWELACTV4'
)

patterns <- c(
  "^RREL", "^ET2REL", "^SNAP", "^TEHC",
  "^EOWN", "^EJSOWN", "^EJOOWN", "^EOOWN",
  "^TJS", "^TJO", "^VAL", "^EVEH", "^TVEH"
)

pattern_vars <- unique(unlist(lapply(patterns, function(p) grep(p, all_cols, value = TRUE))))
keep_vars    <- unique(c(exact_vars, pattern_vars))
keep_vars    <- keep_vars[keep_vars %in% all_cols]

pu <- fread(ds, sep = "|", select = keep_vars)
names(pu) <- toupper(names(pu))

dw <- "data/rw2019.csv"
rw <- fread(dw, sep = "|")
names(rw) <- toupper(names(rw))

sipp2019 <- left_join(pu, rw, by = c("SSUID", "PNUM", "MONTHCODE"))
```

## Filter to July 2019 and generate new SNAP unit ID

Households are uniquely identified in the SIPP by SSUID and SHHADID, but some households may contain multiple SNAP units, defined as people who live and purchase food together. The SIPP does not directly ask about this, but we approximate it based on household relationships. We assume that family members and partners living together are part of the same SNAP unit, while roommates/boarders/other non-family members are not.

```{r}
#| output: false

sipp_0719 <- sipp2019 |>
  filter(RIN_UNIV == 1 & MONTHCODE == 7) |>
  mutate(hhid = paste0(as.character(SSUID), as.character(SHHADID))) |>
  group_by(hhid) |>
  mutate(snap_split = ERELRPE %in% 16:18, # splitting cases that are roommates/boarders/other non-family into separate SNAP unit
         split_id = if_else(snap_split, cumsum(snap_split), NA_integer_),
         snapuid = case_when(!snap_split ~ 1,
         snap_split ~ split_id + 1)
) |> ungroup() |>
  mutate(snap_hhid = paste0(hhid, snapuid))

# Brief checks to make sure split went correctly
sipp_0719 |>
  group_by(hhid) |>
  summarise(n_snap_units = n_distinct(snap_hhid)) |>
  count(n_snap_units)

sipp_0719 |>
  group_by(hhid) |>
  filter(n_distinct(snap_hhid) > 1) |>
  arrange(hhid, snap_hhid) |>
  select(hhid, ERELRPE, snap_hhid) |>
  ungroup() |>
  slice_head(n = 50)

```

## FPL and Deduction Limits

In this section we set the baseline federal poverty limit (FPL),
standard deduction, and maximum shelter deductions set by the USDA FNS.

For each SNAP household, we find the 2019 federal poverty level using
the The U.S. Department of Health and Human Service's published
guidelines, divided by 12 to specifically capture monthly poverty level.
[@2019pov]. We use a vectorized approach for baseline FPL limits, which
then allows us to modify them to set limit simulations later on. While,
Alaska, Guam, and Hawaii have varying FPL limits, to capture their
differential costs of living, given their small sample size in the SIPP,
we did not include their specific FPL's. The standard deduction and
shelter deduction are set by FNS [@fns2018]

```{r}
# 2019 monthly FPL (contiguous 48 states and DC)
fpl_monthly    <- c(1041, 1410, 1778, 2146, 2515, 2883, 3251, 3620)
fpl_additional <- 369

# 2019 standard deductions by household size (caps at 6)
std_ded_tbl <- c(164, 164, 164, 174, 204, 234)

# Excess shelter deduction cap for non-elderly/disabled households
shelter_cap <- 552

get_fpl <- function(n){
  fpl_monthly[pmin(n, 8)] + pmax(n - 8, 0) * fpl_additional
}
get_std_ded <- function(n){
  std_ded_tbl[pmin(n, 6)]
}
```

## Aggregate to Household-Level SNAP Filing Units

The proceeding code in this section collapses the individual to a group
level of analysis (SNAP Household). Then at the household level we create
every variable, usually either a binary flag (TRUE or FALSE) or
continuous limit needed to determine that household's SNAP eligibility.
In this section we do not fully determine whether a household is
eligible just simply flag data needed for determination. We use the 
Congressional Research Service's primer on SNAP eligibility to guide us
in collecting the information we need to determine eligibility and then
making those eligibility determinations [@crs2025].

There are 5 sets of determinations we make at this stage:

1.  **Composition:** number of members of household and whether there is
    someone who is elderly or disabled in the household. Flags for
    citizenship and state of residence are also included.
2.  **Categorical Eligibility:** whether all members of the family
    receive TANF or SSI.
3.  **Income and Assets:** the sum of gross income, earned income,
    countable assets, and vehicle assets.
4.  **Deductions** (case-specific): whether the household had any child
    support payments, child care costs, or out-of-pocket medical
    expenses.
5.  **ABAWD (Able-Bodied Adults without Dependents):** whether the
    household has children, their ages, whether adults works, how old
    they are, whether they are disabled, and how much they work per week. 
    ABAWDs generally may only receive SNAP benefits for 3 months out of a 
    3-year period, unless they are working at least 20 hours per week or 
    participating in training. For simplicity, we assume that any ABAWD 
    not meeting the requirement in July 2019 is ineligible, unless they 
    are in a state where the requirements are fully waived.

SNAP filing units are people who live and purchase food together. Given
that SIPP doesn't account for people who consume food together, we use
SSUID to account for the pseudo-economic unit.

It's important to note the SIPP is very well known for under counting
income information [@meyer2015]. As such we decided for the purposes of
this microsimulation exercise to apply a modest income-multiplier to
account for these instances.

Given that we construct our own SNAP households based on SNAP
methodology (based on those who share food rather than the SIPP
furnished household ID), we use the SIPP User Guide's recommendation to
create household weights by simply taking a reference person's weight to
represent the SNAP household [@smith].

```{r}
# Aggregation
hh_sipp <- sipp_0719 |>
  group_by(snap_hhid) |>
  summarize(
    hh_size = n(),
    has_elderly = any(TAGE >= 60, na.rm = TRUE),
    has_disabled = any(RDIS == 1, na.rm = TRUE),
    has_ssi = all(RSSI_MNYN == 1, na.rm = TRUE),
    has_tanf = all(RTANF_MNYN == 1, na.rm = TRUE),
    # Income
    hh_gross_inc = sum(TPTOTINC, na.rm = TRUE) * 1.25, # sipp undercounts apply this factor
    hh_earn = sum(TPEARN, na.rm = TRUE) * 1.25, # sipp undercounts apply this factor
    # Deductions
    child_sup_ded = sum(TAMOUNTPAID, na.rm = TRUE),
    child_care_ded = sum(TPAYWK, na.rm = TRUE) * 4,    
    medical_costs = sum(TMDPAY, na.rm = TRUE),
    # Shelter (December of reference year as proxy)
    shelter_costs = coalesce(first(na.omit(THVAL_RMU)), 0),
    # Countable assets
    countable_assets = sum(TVAL_BANK, na.rm = TRUE) + sum(TVAL_BOND, na.rm = TRUE) +
                       sum(TVAL_BUS,  na.rm = TRUE) + sum(TVAL_STMF, na.rm = TRUE) +
                       sum(TVAL_RENT, na.rm = TRUE) + sum(TVAL_RE,   na.rm = TRUE) +
                       sum(TVAL_OTH,  na.rm = TRUE),
    vehicle_assets = sum(TVAL_VEH, na.rm = TRUE),
    tveh1_val = first(TVEH1VAL),
    tveh2_val = first(TVEH2VAL),
    # ABAWD flags
    has_children = any(TAGE < 18, na.rm = TRUE),      
    youngest_child_age = if (any(TAGE < 26, na.rm = TRUE)) 
                           as.numeric(min(TAGE[TAGE < 26], na.rm = TRUE))
                         else NA_real_,
    has_abawd_age = any(TAGE >= 18 & TAGE <= 49 & # baseline 18-49 flag
                       (is.na(RDIS) | RDIS != 1), na.rm = TRUE),
    has_adult_nondis_50_59 = any(TAGE >= 50 & TAGE <= 59 &
                                (is.na(RDIS) | RDIS != 1), na.rm = TRUE),
    has_adult_nondis_60_65 = any(TAGE >= 60 & TAGE <= 65 &
                                (is.na(RDIS) | RDIS != 1), na.rm = TRUE),
    max_work_hrs = max(c(0, TMWKHRS[TAGE >= 18 &         
                       (is.na(RDIS) | RDIS != 1)]), na.rm = TRUE),
    has_cit_or_elig_noncit = any(ECITIZEN == 1 | (TIMSTAT == 1 & TYRENTRY <= 2014), na.rm = TRUE), 
    state_fips = first(TEHC_ST),
    weight = first(WPFINWGT[!is.na(ERELRPE) & ERELRPE == min(ERELRPE, na.rm = TRUE)]),
    snap_actual = any(RSNAP_MNYN == 1, na.rm = TRUE)
  )
```

## Describing the Aggregated SNAP Household Data

We examine the distribution of household gross income in the data for
all households, and households actually receiving SNAP

```{r}
#| warning: false
library(scales)
# Limit to households earning less than $60k gross income for aesthetic purposes
inc_cutoff <- hh_sipp |>
  filter(hh_gross_inc <= 60000)
ggplot() +
  # All households earning less than $60k
  geom_histogram(data = inc_cutoff, 
                 mapping = aes(x = hh_gross_inc, weight = weight,
                               fill = "All households"),
                 bins = 100, alpha = 0.5) +
  # Households receiving SNAP
  geom_histogram(data = subset(hh_sipp, snap_actual == TRUE),
                 mapping = aes(x = hh_gross_inc, weight = weight,
                               fill = "Households receiving SNAP"),
                 bins = 100, alpha = 0.5) +
  scale_fill_manual(values = c("All households" = "red",
                      "Households receiving SNAP" = "blue")) +
  scale_y_continuous(labels = label_number(scale_cut = cut_short_scale())) +
  scale_x_continuous(limits = c(0, 60000), 
                     labels = dollar_format(prefix = "$")) +
  labs(title = "Weighted Distribution of Income for Households with <$60k", 
       subtitle = "For all households and households receiving SNAP",
       x = "Household gross income", y = "Number of households",
       fill = "Household") +
  theme_minimal()
```

The income distribution is heavily right skewed. The gross income
distribution of SNAP households has the same shape as for all
households, but the overall distribution is more concentrated on the
left, as we expect. Next we examine the distribution of countable assets
for all households in the data (savings, stocks, money market accounts).

```{r}
#| warning: false
hh_sipp |>
  ggplot() +
  geom_histogram(mapping = aes(x = countable_assets, weight = weight),
                 bins = 100) +
  scale_y_continuous(limits = c(0, 15000000),
                     labels = label_number(scale_cut = cut_short_scale())) +
  scale_x_continuous(limits = c(0, 15000000), 
                     labels = dollar_format(prefix = "$")) +
  labs(title = "Weighted Distribution of Countable Assets of SIPP SNAP Filing Units",
       x = "Countable Assets", y = "Number of households") +
  theme_minimal()
```

The distribution of countable assets is more heavily right skewed, and
with a much longer tail. Next we visualize the number of adults in the
data that meet or don't meet the ABAWD work requirement.

```{r}
abawd_summary <- hh_sipp |>
  mutate(
    abawd_status = case_when(
      has_ssi | has_tanf ~ "Exempt: SSI/TANF",
      !has_abawd_age ~ "Exempt: Age out of range",
      !is.na(youngest_child_age) & youngest_child_age < 18 ~ "Exempt: Has qualifying dependent",
      has_abawd_age & max_work_hrs >= 20 ~ "ABAWD: Meets work requirement",
      has_abawd_age & max_work_hrs < 20 ~ "ABAWD: Does not meet requirement",
      TRUE ~ "Other"
    ),
    top_level = if_else(str_starts(abawd_status, "ABAWD"), "Has ABAWD", "No ABAWD")
  ) |>
  group_by(top_level, abawd_status) |>
  summarize(n_weighted = sum(weight, na.rm = TRUE), .groups = "drop") |>
  mutate(share = n_weighted / sum(n_weighted))
abawd_summary
```

```{r}
# Inner ring - top level
inner <- abawd_summary |>
  group_by(top_level) |>
  summarize(share = sum(share))
# Outer ring - subsections
outer <- abawd_summary
# Colors
colors <- c("Has ABAWD" = "#d6604d",
            "ABAWD: Meets work requirement" = "#db7464",
            "ABAWD: Does not meet requirement" = "#ecb5ac",
            "No ABAWD" = "#2166ac",
            "Exempt: SSI/TANF" = "#92c5de",
            "Exempt: Age out of range" = "#74add1",
            "Exempt: Has qualifying dependent" = "#4393c3")
ggplot() +
  # Inner ring
  geom_col(data = inner,
           aes(x = 1, y = share, fill = top_level),
           width = 0.5) +
  # Outer ring
  geom_col(data = outer,
           aes(x = 2, y = share, fill = abawd_status),
           width = 0.5) +
    scale_fill_manual(values = colors) +
  coord_polar(theta = "y") +
  xlim(0, 2.5) +  
  theme_void() +
  labs(title = "ABAWD Status of SNAP Households",
       fill = NULL)
```

Most SNAP households in our data are exempt from ABAWD requirements,
mainly for being out of the age range. About 1/4 of households have an
adult that would likely qualify for the ABAWD restrictions, and of those
households, the qualification is mostly from meeting the work
requirement.

## Eligibility Function

The goal of this section is to estimate national SNAP household
eligibility under various sets of simulations. Fundamentally this
section accomplishes two things.

First, we code three broad categories of policy baselines into the
function arguments which can then be modified for simulation. These are
income, asset, and ABAWD work policy baselines. Most of these simply
just restate the federal guidelines.

-   **Income**: sets the national gross income limit to 130% of FPL and
    net income limit to 100% of FPL.

    -   Net Income is calculated by subtracting a 20% earned income
        deduction, a standard deduction, child support/ care costs,
        excess shelter deduction (housing above 50% of net income, but
        capped), and medical expenses (elderly and disabled only).
        Additionally it adjust gross limits for states with broad based
        categorical eligibility.

-   **Asset**: keeps each state's and the federal asset limit intact.
    The federal default for assets is \$2250 and vehicle assets are
    counted using state rules.

-   **ABAWD**: The federal requirement for ABAWD requirements are
    non-disabled adults aged 18–49 without dependent children must work
    at least 20 hours per week unless their state has a work exemption.
    HH's that receive SSI or TANF are exempt. The simulation allows the
    ABAWD age ceiling, minimum hours, and dependent child age cutoff to
    be varied as policy parameters.

As such simply applying a data frame to the baseline function, produces
a baseline SNAP household eligibility simulation.

Secondly, using these baseline levels and the argumentation described
within the code to 5 categories of eligibility which need to be met for
SNAP. **Basically they must be income, asset, work, and citizen
eligible. Or they must meet categorically eligibility and be work and
citizen eligible.**

-   Categorical Eligibility: Households where all members receive either
    SSI or TANF automatically pass income and asset tests.

-   Income Eligible (adjusted for BBCE): households must pass both a
    gross income test a net income test as described earlier, but states
    with BBCE adjust these limits.

-   Asset Eligibility: Countable assets including accounts, bonds,
    stocks, vehicles, etc must be below an applicable limit which varies
    by state, although there are federal limits.

-   Work Eligibility: Basically the process described earlier.

-   Citizen Eligibility: At least one householder must be a U.S citizen
    or legal permanent resident who entered the country at least 5 years
    ago.

```{r}
# SNAP Eligibility determination
snap_elig_sim <- function(df,
    # Note all of these can be changed as simulations: Income 
    gross_pct = 1.30,     # gross income limit as fraction of FPL (federal default: 1.30)
    net_pct = 1.00,     # net income limit as fraction of FPL   (federal default: 1.00)
    use_bbce= TRUE,     # apply state BBCE income extensions
    # Asset 
    elim_asset = FALSE,  # TRUE = no asset test for anyone
    asset_limit = NULL,   # set a uniform asset limit in dollars; NULL = federal/BBCE default
    # Work 
    work_req = NULL, # NULL = use state rules; "full_exemption"; "no_exemption"
    abawd_max_age = 49,   # upper age of ABAWD range (federal default: 49)
    min_work_hrs = 20,   # minimum weekly hours to meet work requirement 
    dep_age = 18    # child must be under this age to exempt adult from ABAWD 
) { with(df, 
         {
  # State snap policy
  sp <- snapdb[match(state_fips, snapdb$state_fips), ]

  # FPL thresholds
  fpl       <- get_fpl(hh_size)
  gross_lmt <- fpl * gross_pct
  net_lmt   <- fpl * net_pct
  fed_asset_lmt <- if_else(has_elderly | has_disabled, 3500, 2250)
  if (!is.null(asset_limit)) fed_asset_lmt <- rep(asset_limit, length(hh_size))

  # Net income calculation
  std_ded <- get_std_ded(hh_size)
  earn_ded <- hh_earn * 0.20
  net_pre_shelter <- pmax(hh_gross_inc - std_ded - earn_ded - child_sup_ded - child_care_ded, 0)
  shelter_excess  <- pmax(shelter_costs - 0.50 * net_pre_shelter, 0)
  shelter_ded <- if_else(has_elderly | has_disabled,
                             shelter_excess, pmin(shelter_excess, shelter_cap))
  medical_ded <- if_else(has_elderly | has_disabled, pmax(medical_costs - 35, 0), 0)
  net_inc <- pmax(net_pre_shelter - shelter_ded - medical_ded, 0)

  # Categorical eligibility (SSI/TANF bypasses income and asset tests)
  cat_elig <- has_ssi | has_tanf

  # BBCE — look up from state policy; disabled when use_bbce = FALSE
  bbce_on <- use_bbce & !is.na(sp$bbce) & sp$bbce == 1
  bbce_applies <- bbce_on & !(coalesce(sp$bbce_child, 0L) == 1L & !has_children)

  gross_lmt_bbce <- if_else(bbce_applies & coalesce(sp$bbce_inclmt, -9L) > 0,
                             fpl * (sp$bbce_inclmt / 100), gross_lmt)
  gross_lmt_eld <- if_else(bbce_applies & coalesce(sp$bbce_elddisinclmt, -9L) > 0,
                             fpl * (sp$bbce_elddisinclmt / 100), gross_lmt_bbce)
  gross_lmt_eff <- if_else(has_elderly | has_disabled, gross_lmt_eld, gross_lmt_bbce)

  asset_test_on <- !(bbce_applies & !is.na(sp$bbce_asset) & sp$bbce_asset == 1)
  asset_lmt_eff <- case_when(
    bbce_applies & coalesce(sp$bbce_a_amt, 0) > 0 ~ sp$bbce_a_amt * 1000,
    TRUE ~ fed_asset_lmt
  )
  if (elim_asset) asset_test_on <- rep(FALSE, length(hh_size))

  # Vehicle countable value — look up state vehicle exclusion rules
  vehicle_countable <- case_when(
    sp$vehexclall == 1 ~ 0,
    sp$vehexclone == 1 ~ pmax(vehicle_assets - coalesce(tveh1_val, 0), 0),
    TRUE ~ vehicle_assets
  )
  net_worth <- countable_assets + vehicle_countable

  # Eligibility tests
  gross_elig <- (has_elderly | has_disabled) | hh_gross_inc <= gross_lmt_eff
  net_elig   <- net_inc      <= net_lmt
  asset_elig <- !asset_test_on | cat_elig | (net_worth <= asset_lmt_eff)

  eff_work_req <- if (!is.null(work_req)) rep(work_req, length(hh_size)) else sp$work_req
  has_abawd_age_eff  <- has_abawd_age |
                        (abawd_max_age >= 55 & has_adult_nondis_50_59) |
                        (abawd_max_age >= 60 & has_adult_nondis_60_65)
  has_qualifying_dep <- !is.na(youngest_child_age) & youngest_child_age < dep_age
  meets_work_eff <- max_work_hrs >= min_work_hrs
  has_abawd <- has_abawd_age_eff & !has_qualifying_dep & !has_ssi & !has_tanf
  work_elig <- eff_work_req == "full_exemption" | !has_abawd | meets_work_eff

  citizen_elig <- coalesce(has_cit_or_elig_noncit, TRUE) | cat_elig

  (cat_elig | (gross_elig & net_elig)) & asset_elig & work_elig & citizen_elig
})}

```

## Scenario Functions

Now that we have our function to simulate SNAP eligibility, we create a
function to run different SNAP policy scenarios, which will allow us to
estimate how eligibility rates change as aspects of SNAP policy are
adjusted. We also write a function to summarize scenarios and calculate
the number of eligible SNAP households under different policies.

```{r}
run_scenario <- function(data, scenario_name, ...) {
  data |>
    mutate(snap_eligible = snap_elig_sim(pick(everything()), ...),
           scenario = scenario_name)
}

summarize_scenario <- function(data) {
  data |>
    ungroup() |>
    summarize(
      scenario = first(scenario),
      n_eligible_weighted = sum(snap_eligible * weight, na.rm = TRUE),
      n_total_weighted = sum(weight, na.rm = TRUE),
      pct_eligible= round(sum(snap_eligible * weight, na.rm = TRUE) /
                           sum(weight, na.rm = TRUE) * 100, 1))
}
```

## Run Scenarios at National Level

Roughly participated in snap in 2019 [@reaching]. July 2019 had 19
million households [@snapdat]. As such assuming the SIPP properly
captures income, asset information, then we should have roughly 23.5
million households.

Here, we start by running our baseline scenario for the year and month
we have chosen (July 2019). This is our baseline for calculating percent
changes in eligibility under the other scenarios.

**Income.** We have two policy scenarios that adjust how much income is
needed to qualify for SNAP. First, we change the gross income limit to
200% of the federal poverty line (FPL). Then, remove broad-based
categorical eligibility (BBCE).

**Assets.** We have two scenarios that adjust asset limits for SNAP
eligibility: eliminating the asset test, and raising the asset limit to
\$10,000.

**Work requirements.** There are several work requirements for SNAP
eligibility, and we estimate how rates would change if the following
policies were adopted with respect to work requirements: waive ABAWD,
fully enforce ABAWD, require a minimum of 30 hours/week, make the
dependent age cutoff 5 years old, and apply ABAWD to individuals aged
18-65.

```{r}
# Baseline: 2019 state rules as-is
baseline <- run_scenario(hh_sipp, "Baseline (2019)")

# Income
income_200pct  <- run_scenario(hh_sipp, "Gross limit 200% FPL",
                               gross_pct = 2.00)

income_no_bbce <- run_scenario(hh_sipp, "Federal rules only (no BBCE)",
                               use_bbce = FALSE)

# Assets
asset_elim <- run_scenario(hh_sipp, "Eliminate asset test",
                           elim_asset = TRUE)

asset_10k  <- run_scenario(hh_sipp, "Raise asset limit to $10k",
                           asset_limit = 10000)

# Work requirement scenarios 
work_waive <- run_scenario(hh_sipp, "Waive ABAWD nationally",
                             work_req = "full_exemption")

work_enforce <- run_scenario(hh_sipp, "Enforce ABAWD nationally",
                             work_req = "no_exemption")

work_30hrs <- run_scenario(hh_sipp, "30-hour requirement", min_work_hrs = 30)

work_dep5    <- run_scenario(hh_sipp, "Dep age cutoff 5", dep_age = 5)

work_age65 <- run_scenario(hh_sipp, "ABAWD age 18-65", abawd_max_age = 65)
```

## National Results

```{r}
results <- bind_rows(
  summarize_scenario(baseline),
  summarize_scenario(income_200pct),
  summarize_scenario(income_no_bbce),
  summarize_scenario(asset_elim),
  summarize_scenario(asset_10k),
  summarize_scenario(work_waive),
  summarize_scenario(work_enforce),
  summarize_scenario(work_30hrs),
  summarize_scenario(work_dep5),
  summarize_scenario(work_age65),
) |>
  mutate(
    change_from_baseline = n_eligible_weighted - n_eligible_weighted[scenario == "Baseline (2019)"],
    pct_change = round(100 * change_from_baseline / n_eligible_weighted[scenario == "Baseline (2019)"], 1)
  )

results
```

```{r}
results |>
  ggplot(aes(reorder(scenario, pct_eligible), pct_eligible)) +
  geom_col() +
  geom_text(aes(label = paste0(pct_eligible, "%")), hjust = -0.1) +
  coord_flip() +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.15)),
    labels = scales::label_percent(scale = 1)) +
  labs(
    title = "SNAP Eligibility Rate by Policy Scenario",
    x = NULL,
    y = "% Eligible") +
  theme_minimal()
```

For baseline eligibility in July 2019, we estimate that approximately 26
million households are eligible for SNAP. This is similar to the 23.5
million that we expected. Adjusting the gross income limit to 200% of
the FPL minimally increases the number of eligible households by 0.5
percent from baseline, while enforcing ABAWD nationally decreases the
eligibility rate by 0.5 percent. The greatest estimated changed from
baseline is eliminating categorical eligibility and applying only
federal eligibility rules nationally. We estimate that the number of
eligible households would decrease to about 17 million households, a
decrease of 35.4 percent. Eliminating the asset test and waiving ABAWD
nationally are both estimated to increase eligibility rates by about 13
percent, to about 29.6 million eligible households.

### Baseline Income Distribution of SNAP eligibility

**Income Decile:**

First we want to understand the underlying income distribution of SNAP
households to be able to model how this changes under various policy
specifications. We first look at income decile eligibility distribution.

```{r}
distributional_table <- baseline |>
  mutate(inc_decile = ntile(hh_gross_inc, n = 10)) |>
  group_by(inc_decile) |>
  summarize(
    min_income = min(hh_gross_inc),
    max_income = max(hh_gross_inc),
    pct_eligible = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1),
    n_eligible_weighted = sum(snap_eligible * weight, na.rm = TRUE)
  )

distributional_table
```

```{r}
distributional_table |>
  ggplot(aes(inc_decile, pct_eligible)) +
  geom_col() +
  geom_text(aes(label = paste0(pct_eligible, "%")), vjust = -1) +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.1)),
    labels = scales::label_percent(scale = 1)) +
  labs(
    title = "SNAP Eligibility Rate by Household Income Decile (Baseline 2019)",
    x = "Household Income Decile",
    y = "% Eligible") +
  theme_minimal()

```

It'll be difficult to demonstrate a change for every income decile,
however in our analysis we can see that at least half in the bottom 20%
of the distribution are eligible for SNAP and at least 10% of every
decile in the bottom 50% is eligible for SNAP. As such the quartile
distribution may be a better from of reference

**Quartile Distribution:**

```{r}
quartile_distribution <- baseline |>
  mutate(inc_quartile = ntile(hh_gross_inc, n = 4)) |>
  group_by(inc_quartile) |>
  summarize(
    min_income = min(hh_gross_inc),
    max_income = max(hh_gross_inc),
    pct_eligible = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1),
    n_eligible_weighted = sum(snap_eligible * weight, na.rm = TRUE)) |>
  ggplot(aes(inc_quartile, pct_eligible)) +
  geom_col() +
  geom_text(aes(label = paste0(pct_eligible, "%")), vjust = -1) +
  scale_x_continuous(breaks = 1:4) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.1)),
    labels = scales::label_percent(scale = 1)) +
  labs(
    title = "SNAP Eligibility Rate by Household Income Quartile (Baseline 2019)",
    x = "Household Income Quartile",
    y = "% Eligible") +
  theme_minimal()

quartile_distribution
```

This suggests that the bottom quarter is most directly impacted by our
SNAP eligibility formula. However, at least 15% of the next quartile is
also impacted. We take the bottom quarter of the population and model
how their eligibility changes nationally under the different simulation.

### Income Distribution of SNAP eligibility by Simulated Scenario

**Bottom Quartile by Simulation**

```{r}
bind_rows(baseline, income_200pct, income_no_bbce, asset_elim, asset_10k,
          work_waive, work_enforce, work_30hrs, work_dep5, work_age65) |>
  group_by(scenario) |>
  mutate(inc_quartile = ntile(hh_gross_inc, n = 4)) |>
  filter(inc_quartile == 1) |>
  summarize(
    pct_eligible = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1)
  ) |>
  ggplot(aes(reorder(scenario, pct_eligible), pct_eligible)) +
  geom_col() +
  geom_text(aes(label = paste0(pct_eligible, "%")), hjust = -0.1) +
  coord_flip() +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.15)),
    labels = scales::label_percent(scale = 1)) +
  labs(
    title = "SNAP Eligibility in Bottom Income Quartile by Scenario",
    x = NULL,
    y = "% Eligible") +
  theme_minimal()
```

**Second Bottom Quartile by Simulation**

```{r}
bind_rows(baseline, income_200pct, income_no_bbce, asset_elim, asset_10k,
          work_waive, work_enforce, work_30hrs, work_dep5, work_age65) |>
  group_by(scenario) |>
  mutate(inc_quartile = ntile(hh_gross_inc, n = 4)) |>
  filter(inc_quartile == 2) |>
  summarize(
    pct_eligible = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1)) |>
  ggplot(aes(reorder(scenario, pct_eligible), pct_eligible)) +
  geom_col() +
  geom_text(aes(label = paste0(pct_eligible, "%")), hjust = -0.1) +
  coord_flip() +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.15)),
    labels = scales::label_percent(scale = 1)) +
  labs(
    title = "SNAP Eligibility in Second Income Quartile by Scenario",
    x = NULL,
    y = "% Eligible") +
  theme_minimal()
```

Across the bottom half of the income distribution patterns emerge:

1.  BBCE helps ensure lower income families can have access to food
    assistance and are the single biggest driver SNAP participation in
    our models.
2.  OBBBA style reforms (ABAWD work requirements) are the most
    regressive and lower families SNAP eligibility across the board.
    1.  **In the lowest quartile simply expanding the ABAWD age range
        from up to 49 to 65 means 5% of families in the lowest quartile
        lose SNAP.**

**We can also just for the sake of depth look at the distributions for
each policy by quartile**

```{r}
bind_rows(baseline, income_200pct, income_no_bbce, asset_elim, asset_10k,
          work_waive, work_enforce, work_30hrs, work_dep5, work_age65) |>
  group_by(scenario) |>
  mutate(inc_quartile = ntile(hh_gross_inc, n = 4)) |>
  group_by(scenario, inc_quartile) |>
  summarize(
    pct_eligible = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1),
    .groups = "drop") |>
  ggplot(aes(x = factor(inc_quartile), y = reorder(scenario, pct_eligible), fill = pct_eligible)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(pct_eligible, "%")), size = 3.2) +
  scale_fill_gradient(low = "white", high = "#2166ac",
                      name = "% Eligible") +
  labs(
    title = "SNAP Eligibility Rate by Income Quartile and Scenario",
    x = "Income Quartile",
    y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid = element_blank())
```

This generally doesn't reveal anything too stark about SNAP however it
does indicate that BBCE policies also affect those who aren't
necessarily in the bottom half of the distribution.

### Pseudo OBBBA ABAWD Work Requirement Simulation

All of the prior simulations only toggle one scenario at a time.
However, if we edit the scenarios by raising the age of exemption to 65
and edit the dependent child clause to not exempt families where the
youngest child is over 14 then we can somewhat model OBBBA. We can't
necessarily cover some of the provisions like stricter work waivers or
cost share behavorial responses however.

```{r}
obbba_simulation <- run_scenario(hh_sipp, "OBBBA (approx.)",
                                 abawd_max_age = 65,
                                 dep_age = 14)
```

```{r}
bind_rows(
  summarize_scenario(baseline),
  summarize_scenario(obbba_simulation)) |>
  ggplot(aes(reorder(scenario, pct_eligible), pct_eligible)) +
  geom_col() +
  geom_text(aes(label = paste0(pct_eligible, "%")), vjust = -0.5) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.15)),
    labels = scales::label_percent(scale = 1)) +
  labs(
    title = "SNAP Eligibility: Baseline vs. OBBBA Work Requirement Changes",
    x = NULL,
    y = "% Eligible") +
  theme_minimal()
```

## State-level eligibility results

Because states have different baseline eligibility rates, and baseline
SNAP eligibility rules, we are also interested in summarizing how these
different policy changes would change eligibility rates at the state
level. We write a function to summarize the estimated number of eligible
households for each state.

```{r}
summarize_by_state <- function(data) {
  data |>
    group_by(state_fips, scenario) |>
    summarize(
      n_eligible_weighted = sum(snap_eligible * weight, na.rm = TRUE),
      n_total_weighted    = sum(weight, na.rm = TRUE),
      pct_eligible = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1),
      .groups = "drop"
    )
}
```

Next, we generate the same policy scenarios described above, but
summarized for their impact on each state's eligibility rates.

```{r}
# Baseline
baseline_state <- hh_sipp |>
  run_scenario("Baseline") |>
  summarize_by_state()

# Income
twohundred_state  <- hh_sipp |>
  run_scenario("Gross limit 200% FPL", gross_pct = 2.00) |>
  summarize_by_state()

no_bbce_state <- hh_sipp |>
  run_scenario("Federal rules only (no BBCE)", use_bbce = FALSE) |>
  summarize_by_state()

# Assets
asset_elim_state <- hh_sipp |>
  run_scenario("Eliminate asset test", elim_asset = TRUE) |>
  summarize_by_state()

asset_10k_state  <- hh_sipp |>
  run_scenario("Raise asset limit to $10k", asset_limit = 10000) |>
  summarize_by_state()

# Work requirement scenarios 
work_waive_state <- hh_sipp |>
  run_scenario("Waive ABAWD nationally", work_req = "full_exemption") |>
  summarize_by_state()

work_enforce_state <- hh_sipp |>
  run_scenario("Enforce ABAWD nationally", work_req = "no_exemption") |>
  summarize_by_state()

work_30hrs_state <- hh_sipp |>
  run_scenario("30-hour requirement", min_work_hrs = 30) |>
  summarize_by_state()

work_dep5_state <- hh_sipp |>
  run_scenario("Dep age cutoff 5", dep_age = 5) |>
  summarize_by_state()

work_age65_state <- hh_sipp |>
  run_scenario("ABAWD age 18-65", abawd_max_age = 65) |>
  summarize_by_state()
```

## State Level Results

```{r}
state_results <- bind_rows(baseline_state, twohundred_state,
                           no_bbce_state, asset_elim_state, asset_10k_state,
                           work_waive_state, work_enforce_state, work_30hrs_state,
                           work_dep5_state, work_age65_state) |>
  group_by(state_fips) |>
  mutate(change_from_baseline = n_eligible_weighted - n_eligible_weighted[scenario == "Baseline"],
         pct_change = round(100 * change_from_baseline / n_eligible_weighted[scenario == "Baseline"],1)) |>
  ungroup() |>
  arrange(state_fips)

state_results
```

# Visualize State Level Results

```{r}
visualize_scenario <- function(df, scen) {
  # State geometry from tigris
  states <- states(cb = TRUE, resolution = "20m") |>
    st_transform(5070) |>
    shift_geometry()
  
  # Separate out small states that will need custom labels
  small_states <- c("VT","NH","MA","RI","CT","NJ","DE","MD","DC")
  
  # Merge state geometry with eligibility microsim data
  plot_data <- df |> 
    filter(scenario == scen) |>
    mutate(STATEFP = str_pad(state_fips, width = 2, side = "left", pad = "0")) |>
    left_join(states, by = "STATEFP") |>
    st_as_sf() |>
    mutate(is_small = STUSPS %in% small_states)
  
  # Label positions and table for small states
  centroids <- st_centroid(plot_data)
  
  coords <- centroids %>%
    mutate(
      x = st_coordinates(.)[,1],
      y = st_coordinates(.)[,2]
    ) %>%
    st_drop_geometry()
  
  table_positions <- coords |>
    filter(is_small) |>
    arrange(desc(y)) |> 
    mutate(
      table_x = max(coords$x) + 6e5,
      table_y = seq(max(coords$y), length.out = n(), by = -2e5)
    )
  
  coords <- coords |>
    left_join(table_positions |> select(STUSPS, table_x, table_y), by = "STUSPS")
  
  # Make plot
  ggplot(plot_data) +
    geom_sf(aes(fill = pct_eligible), color = "white") +
    # Labels for large states
    geom_text(
      data = coords |> filter(!is_small),
      aes(x = x, y = y, label = pct_eligible),
      size = 3
    ) +
    # Segments for small states
    geom_segment(
      data = coords |> filter(is_small),
      aes(x = x, y = y, xend = table_x, yend = table_y),
      color = "gray50"
    ) +
    # Separate labels for small states
    geom_text(
      data = coords |> filter(is_small),
      aes(x = table_x, y = table_y, 
          label = paste0(STUSPS, ": ", pct_eligible)),
      hjust = 0,
      size = 3
    ) +
    scale_fill_gradient(low = "#BAE1FF", high = "#0B538A") +
    coord_sf(clip = "off") +
    labs(title = scen, fill = "Percent of households SNAP eligible") +
    theme_void() +
    theme(
      plot.title = element_text(hjust = 0.5),
      plot.margin = margin(10, 100, 10, 10),
      legend.position = "bottom"
    )
    
}


```

```{r}
#| results: 'hide'

#| warning: false
#| message: false
visualize_scenario(state_results, scen = "Baseline")

visualize_scenario(state_results, scen = "Gross limit 200% FPL")

visualize_scenario(state_results, scen = "Federal rules only (no BBCE)")

visualize_scenario(state_results, scen = "Eliminate asset test")

visualize_scenario(state_results, scen = "Raise asset limit to $10k")

visualize_scenario(state_results, scen = "Waive ABAWD nationally")

visualize_scenario(state_results, scen = "Enforce ABAWD nationally")

visualize_scenario(state_results, scen = "30-hour requirement")

visualize_scenario(state_results, scen = "Dep age cutoff 5")

visualize_scenario(state_results, scen = "ABAWD age 18-65")

```
