---
title: "SNAP Eligibility Microsimulation"
format:
  html:
    embed-resources: true
editor:
  markdown:
    wrap: 72
execute:
  cache: true
  cache.lazy: false
---

## Overview

This file is fully self-contained. It loads all data from scratch,
builds household-level SNAP filing units, defines a vectorized
eligibility function, and runs baseline and counterfactual policy
scenarios.

## Libraries

```{r}
#| cache: false
library(tidyverse)
library(readxl)
library(data.table)
library(bit64)
library(skimr)
```

## Load SNAP Policy Database
This file contains state by state information on 

```{r}
snapdb <- read_xlsx("data/SNAPPolicyDatabase.xlsx", sheet = "SNAP Policy Database") |>
  filter(yearmonth == 201907) |>
  select(starts_with("state"), starts_with("bbce"), starts_with("veh"))

# Work requirement tiers based on FY2019 Q4 ABAWD waiver status
state_exempt <- c("Alaska", "District of Columbia", "Louisiana", "New Mexico")

state_nonexempt <- c("Alabama", "Arkansas", "Delaware", "Florida", "Indiana",
                     "Iowa", "Kansas", "Maine", "Mississippi", "Missouri",
                     "Nebraska", "North Carolina", "Oklahoma", "South Carolina",
                     "Texas", "Wisconsin", "Wyoming")

snapdb <- snapdb |>
  mutate(work_req = case_when(
    statename %in% state_exempt ~ "full_exemption",
    statename %in% state_nonexempt ~ "no_exemption",
    TRUE ~ "partial_exemption"))
```

## Load SIPP 2019

```{r}
#| cache: false
ds <- "data/pu2019.csv"

all_cols <- names(fread(ds, sep = "|", nrows = 0))

exact_vars <- c(
  # ID variables
  'SSUID', 'PNUM', 'MONTHCODE', 'ERESIDENCEID', 'ERELRPE', 'SPANEL', 'SWAVE',
  'SHHADID', 'THHLDSTATUS', 'RIN_UNIV',

  # Base weight
  'WPFINWGT',

  # State of residence
  'TEHC_ST', 'TST_INTV',

  # Demographics
  'ESEX', 'TAGE', 'TAGE_EHC', 'ERACE', 'EORIGIN', 'EEDUC',
  'ECITIZEN', 'ENATIVITY', 'EIMM_STAT',
  'EDISABL', 'RDIS', 'RDIS_ALT',

  # Household structure
  'TLIVQTR', 'ETENURE', 'EMS',

  # Income
  'TPTOTINC', 'TPEARN', 'TFTOTINC',
  'TUNEMP_AMT', 'TSSI_AMT', 'TSSRX', 'TSSXAMT', 'TTANF_AMT',

  # Deductions
  'TAMOUNTPAID', 'TPAYWK', 'TRENTMORT', 'TUTILS', 'THVAL_RMU',
  'TCHCARE_AMT', 'TMDPAY',

  # Assets (individual-level; shared assets divided equally)
  'TVAL_AST', 'TVAL_BANK', 'TVAL_BOND', 'TVAL_BUS', 'TVAL_ESAV',
  'TVAL_HOME', 'TVAL_RENT', 'TVAL_RE', 'TVAL_RET',
  'TVAL_STMF', 'TVAL_VEH', 'TVAL_OTH',

  # Program participation
  'RFSCOV', 'RSNAP_MNYN', 'RTANF_MNYN', 'RSSI_MNYN', 'RWICYN',

  # ABAWD work requirement variables
  'TMWKHRS', 'EWELACTV2_1', 'EWELACTV2_2', 'EWELACTV2_3', 'EWELACTV4'
)

patterns <- c(
  "^RREL", "^ET2REL", "^SNAP", "^TEHC",
  "^EOWN", "^EJSOWN", "^EJOOWN", "^EOOWN",
  "^TJS", "^TJO", "^VAL", "^EVEH", "^TVEH"
)

pattern_vars <- unique(unlist(lapply(patterns, function(p) grep(p, all_cols, value = TRUE))))
keep_vars    <- unique(c(exact_vars, pattern_vars))
keep_vars    <- keep_vars[keep_vars %in% all_cols]

pu <- fread(ds, sep = "|", select = keep_vars)
names(pu) <- toupper(names(pu))

dw <- "data/rw2019.csv"
rw <- fread(dw, sep = "|")
names(rw) <- toupper(names(rw))

sipp2019 <- left_join(pu, rw, by = c("SSUID", "PNUM", "MONTHCODE"))
```

## Filter to July 2019 and generate new SNAP unit ID
Households are uniquely identified by SSUID and SHHADID, but some households may contain multiple SNAP units, defined as people who live and purchase food together. The SIPP does not directly ask about this, but we approximate it based on household relationships. We assume that family members and partners living together are part of the same SNAP unit, while roommates/boarders/other non-family members are not.

```{r}
sipp_0719 <- sipp2019 |>
  filter(RIN_UNIV == 1 & MONTHCODE == 7) |>
  mutate(hhid = paste0(as.character(SSUID), as.character(SHHADID))) |>
  group_by(hhid) |>
  mutate(snap_split = ERELRPE %in% 16:18, # splitting cases that are roommates/boarders/other non-family into separate SNAP unit
         split_id = if_else(snap_split, cumsum(snap_split), NA_integer_),
         snapuid = case_when(!snap_split ~ 1,
         snap_split ~ split_id + 1)
) |> ungroup() |>
  mutate(snap_hhid = paste0(hhid, snapuid))

# Brief checks to make sure split went correctly
sipp_0719 |>
  group_by(hhid) |>
  summarise(n_snap_units = n_distinct(snap_hhid)) |>
  count(n_snap_units)

sipp_0719 |>
  group_by(hhid) |>
  filter(n_distinct(snap_hhid) > 1) |>
  arrange(hhid, snap_hhid) |>
  select(hhid, ERELRPE, snap_hhid) |>
  ungroup() |>
  slice_head(n = 50)

```

## FPL and Deduction Limits

```{r}
# 2019 monthly FPL (contiguous 48 states and DC)
# https://aspe.hhs.gov/topics/poverty-economic-mobility/poverty-guidelines
fpl_monthly    <- c(1041, 1410, 1778, 2146, 2515, 2883, 3251, 3620)
fpl_additional <- 369

# 2019 standard deductions by household size (caps at 6)
# https://cvlas.org/wp-content/uploads/2022/08/supp-nutrition-assist-prog-SNAP-2019.pdf
std_ded_tbl <- c(167, 167, 167, 178, 208, 238)

# Excess shelter deduction cap for non-elderly/disabled households
shelter_cap <- 552

get_fpl <- function(n){
  fpl_monthly[pmin(n, 8)] + pmax(n - 8, 0) * fpl_additional
}
get_std_ded <- function(n){
  std_ded_tbl[pmin(n, 6)]
}
```

## Aggregate to Household-Level SNAP Filing Units

```{r}
# Aggregation
hh_sipp <- sipp_0719 |>
  group_by(snap_hhid) |>
  summarize(
    hh_size = n(),
    # Flags for exemptions
    has_elderly = any(TAGE >= 60, na.rm = TRUE),
    has_disabled = any(RDIS == 1, na.rm = TRUE),
    has_ssi = any(RSSI_MNYN == 1, na.rm = TRUE),
    has_tanf = any(RTANF_MNYN == 1, na.rm = TRUE),
    # Income
    hh_gross_inc = sum(TPTOTINC, na.rm = TRUE),
    hh_earn = sum(TPEARN, na.rm = TRUE),
    # Deductions
    child_sup_ded = sum(TAMOUNTPAID, na.rm = TRUE),
    child_care_ded = first(TPAYWK[!is.na(TPAYWK)], default = 0) * 4,
    medical_costs = sum(TMDPAY, na.rm = TRUE),
    # Shelter (December of reference year as proxy)
    shelter_costs = coalesce(first(na.omit(THVAL_RMU)), 0),
    # Countable assets — exclude home, retirement, educational savings
    countable_assets = sum(TVAL_BANK, na.rm = TRUE) + sum(TVAL_BOND, na.rm = TRUE) +
                       sum(TVAL_BUS,  na.rm = TRUE) + sum(TVAL_STMF, na.rm = TRUE) +
                       sum(TVAL_RENT, na.rm = TRUE) + sum(TVAL_RE,   na.rm = TRUE) +
                       sum(TVAL_OTH,  na.rm = TRUE),
    vehicle_assets = sum(TVAL_VEH, na.rm = TRUE),
    # Household-level vehicle values for state exclusion rules
    tveh1_val = first(TVEH1VAL),
    tveh2_val = first(TVEH2VAL),
    tveh3_val = first(TVEH3VAL),
    # ABAWD flags — computed at person level, aggregated to household
    has_children = any(TAGE < 18, na.rm = TRUE),
    has_abawd_age = any(TAGE >= 18 & TAGE <= 49 & RDIS != 1, na.rm = TRUE),
    abawd_meets_work = any(
      (TAGE >= 18 & TAGE <= 49 & RDIS != 1) & (TMWKHRS >= 20), na.rm = TRUE),
    has_any_citizen = any(ECITIZEN %in% c(1L, 2L), na.rm = TRUE), # confirm this
    state_fips = first(TEHC_ST),
    weight = first(WPFINWGT),
    snap_actual = any(RSNAP_MNYN == 1, na.rm = TRUE)
  ) |>
  # joining in snap eligibility rules 
  left_join(snapdb, by = "state_fips")

```

## Eligibility Function

Set up to allow us to change baslines, modify policies. Unforuntately we need all of these components in the function. 

```{r}
snap_elig_sim <- function(
    hh_size, hh_gross_inc, hh_earn,
    countable_assets, vehicle_assets, tveh1_val,
    has_elderly, has_disabled, has_ssi, has_tanf,
    has_children, has_abawd_age, abawd_meets_work,
    has_any_citizen,
    shelter_costs, child_sup_ded, child_care_ded, medical_costs,
    bbce, bbce_inclmt, bbce_asset, bbce_a_amt,
    bbce_child, bbce_elddisinclmt,
    vehexclall, vehexclone, work_req,
    # We can add in more baseline uniform scenarios here to change later
    override_gross_pct = 1.30,   # federal default: 130% FPL
    override_net_pct = 1.00,   # federal default: 100% FPL
    override_asset_lmt = NULL,   #use federal/BBCE limits
    override_elim_asset = FALSE,  # TRUE = eliminate asset test nationally
    override_work_req = NULL   # use state-level work_req

    ) 
{
  
  # FPL thresholds
  fpl <- get_fpl(hh_size)
  gross_limit <- fpl * override_gross_pct
  net_limit <- fpl * override_net_pct
  asset_limit <- if_else(has_elderly | has_disabled, 4500, 3000)
  if (!is.null(override_asset_lmt)) asset_limit <- rep(override_asset_lmt, length(hh_size))

  # Net income
  std_ded  <- get_std_ded(hh_size)
  earn_ded <- hh_earn * 0.20
  net_pre_shelter <- pmax(hh_gross_inc - std_ded - earn_ded - child_sup_ded - child_care_ded, 0)
  shelter_excess <- pmax(shelter_costs - 0.50 * net_pre_shelter, 0)
  shelter_ded <- if_else(has_elderly | has_disabled,
                         shelter_excess, pmin(shelter_excess, shelter_cap))
  medical_ded <- if_else(has_elderly | has_disabled, pmax(medical_costs - 35, 0), 0)
  net_inc <- pmax(net_pre_shelter - shelter_ded - medical_ded, 0)

  # Categorical eligibility
  cat_elig <- has_ssi | has_tanf

  # BBCE: only applies to childless households if bbce_child == 0
  bbce_on      <- !is.na(bbce) & bbce == 1
  bbce_applies <- bbce_on & !(coalesce(bbce_child, 0L) == 1L & !has_children)

  gross_limit_bbce <- if_else(bbce_applies & coalesce(bbce_inclmt, -9L) > 0,
                               fpl * (bbce_inclmt / 100), gross_limit)
  gross_limit_eld  <- if_else(bbce_applies & coalesce(bbce_elddisinclmt, -9L) > 0,
                               fpl * (bbce_elddisinclmt / 100), gross_limit_bbce)
  gross_limit_eff  <- if_else(has_elderly | has_disabled, gross_limit_eld, gross_limit_bbce)

  asset_test_on   <- !(bbce_applies & !is.na(bbce_asset) & bbce_asset == 1)
  asset_limit_eff <- case_when(
    bbce_applies & coalesce(bbce_a_amt, 0) > 0 ~ bbce_a_amt * 1000,
    TRUE                                        ~ asset_limit
  )
  if (override_elim_asset) asset_test_on <- rep(FALSE, length(hh_size))

  # Vehicle countable value
  vehicle_countable <- case_when(
    vehexclall == 1 ~ 0,
    vehexclone == 1 ~ pmax(vehicle_assets - coalesce(tveh1_val, 0), 0),
    TRUE            ~ vehicle_assets
  )
  net_worth_countable <- countable_assets + vehicle_countable

  # Individual tests
  gross_elig <- hh_gross_inc <= gross_limit_eff
  net_elig   <- net_inc      <= net_limit
  asset_elig <- !asset_test_on | cat_elig | (net_worth_countable <= asset_limit_eff)

  # Work requirements
  eff_work_req <- if (!is.null(override_work_req)) rep(override_work_req, length(hh_size)) else work_req
  has_abawd    <- has_abawd_age & !has_children & !has_ssi & !has_tanf
  work_elig    <- eff_work_req == "full_exemption" | !has_abawd | abawd_meets_work

  citizen_elig <- coalesce(has_any_citizen, TRUE) | cat_elig

  # Final eligibility
  (cat_elig | (gross_elig & net_elig)) & asset_elig & work_elig & citizen_elig
}
```


## Random Example to determine eligibility
```{r}
snap_elig_sim(
    hh_size = 3, hh_gross_inc = 100, hh_earn = 100,
    countable_assets = 0, vehicle_assets = 0, tveh1_val = 0,
    has_elderly = FALSE, has_disabled = FALSE, has_ssi = FALSE, has_tanf = FALSE,
    has_children = TRUE, has_abawd_age = TRUE, abawd_meets_work = FALSE,
    has_any_citizen = TRUE,
    shelter_costs = 100, child_sup_ded = 0, child_care_ded = 0, medical_costs = 0,
    bbce = -9, bbce_inclmt = -9, bbce_asset = -9, bbce_a_amt = -9,
    bbce_child = 0, bbce_elddisinclmt = -9,
    vehexclall = 0, vehexclone = 0, work_req = "no_exemption")
```
## Scenario Sample

```{r}
run_scenario <- function(data, scenario_name, ...) {
  data |>
    mutate(
      snap_eligible = snap_elig_sim(
        hh_size, hh_gross_inc, hh_earn,
        countable_assets, vehicle_assets, tveh1_val,
        has_elderly, has_disabled, has_ssi, has_tanf,
        has_children, has_abawd_age, abawd_meets_work,
        has_any_citizen,
        shelter_costs, child_sup_ded, child_care_ded, medical_costs,
        bbce, bbce_inclmt, bbce_asset, bbce_a_amt,
        bbce_child, bbce_elddisinclmt,
        vehexclall, vehexclone, work_req,
        ...
      ),
      scenario = scenario_name
    )
}

summarize_scenario <- function(data) {
  data |>
    ungroup() |>
    summarize(
      scenario            = first(scenario),
      n_eligible_weighted = sum(snap_eligible * weight, na.rm = TRUE),
      n_total_weighted    = sum(weight, na.rm = TRUE),
      pct_eligible        = round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1)
    )
}
```

## Run Scenarios

```{r}
baseline <- run_scenario(hh_sipp, "Baseline (2019)")
```
Scenarios we can run
```{r}
scenario1 <- run_scenario(hh_sipp, "Eliminate asset test",
                   override_elim_asset = TRUE)

scenario2 <- run_scenario(hh_sipp, "Gross limit 200% FPL",
                   override_gross_pct = 2.00)

scenario3 <- run_scenario(hh_sipp, "Waive ABAWD nationally",
                   override_work_req = "full_exemption")


```

## Results

```{r}
results <- bind_rows(
  summarize_scenario(baseline),
  summarize_scenario(scenario1),
  summarize_scenario(scenario2),
  summarize_scenario(scenario3))

results
```

