---
title: "SNAP Eligibility Microsimulation"
format:
  html:
    embed-resources: true
editor:
  markdown:
    wrap: 72
execute:
  cache: true
  cache.lazy: false
---

## Overview

This file is fully self-contained. It loads all data from scratch,
builds household-level SNAP filing units, defines a vectorized
eligibility function, and runs baseline and counterfactual policy
scenarios.

## Libraries

```{r}
#| cache: false
library(tidyverse)
library(readxl)
library(data.table)
library(bit64)
library(skimr)
```

## Load SNAP Policy Database

This file contains state by state information on

```{r}
snapdb <- read_xlsx("data/SNAPPolicyDatabase.xlsx", sheet = "SNAP Policy Database") |>
  filter(yearmonth == 201907) |>
  select(starts_with("state"), starts_with("bbce"), starts_with("veh"))

# Work requirement tiers based on FY2019 Q4 ABAWD waiver status
state_exempt <- c("Alaska", "District of Columbia", "Louisiana", "New Mexico")

state_nonexempt <- c("Alabama", "Arkansas", "Delaware", "Florida", "Indiana",
                     "Iowa", "Kansas", "Maine", "Mississippi", "Missouri",
                     "Nebraska", "North Carolina", "Oklahoma", "South Carolina",
                     "Texas", "Wisconsin", "Wyoming")

snapdb <- snapdb |>
  mutate(work_req = case_when(
    statename %in% state_exempt ~ "full_exemption",
    statename %in% state_nonexempt ~ "no_exemption",
    TRUE ~ "partial_exemption"))
```

## Load SIPP 2019

```{r}
#| cache: false
ds <- "data/pu2019.csv"

all_cols <- names(fread(ds, sep = "|", nrows = 0))

exact_vars <- c(
  # ID variables
  'SSUID', 'PNUM', 'MONTHCODE', 'ERESIDENCEID', 'ERELRPE', 'SPANEL', 'SWAVE',
  'SHHADID', 'THHLDSTATUS', 'RIN_UNIV',

  # Base weight
  'WPFINWGT',

  # State of residence
  'TEHC_ST', 'TST_INTV',

  # Demographics
  'ESEX', 'TAGE', 'TAGE_EHC', 'ERACE', 'EORIGIN', 'EEDUC',
  'ECITIZEN', 'ENATIVITY', 'EIMM_STAT',
  'EDISABL', 'RDIS', 'RDIS_ALT',

  # Household structure
  'TLIVQTR', 'ETENURE', 'EMS',

  # Income
  'TPTOTINC', 'TPEARN', 'TFTOTINC',
  'TUNEMP_AMT', 'TSSI_AMT', 'TSSRX', 'TSSXAMT', 'TTANF_AMT',

  # Deductions
  'TAMOUNTPAID', 'TPAYWK', 'TRENTMORT', 'TUTILS', 'THVAL_RMU',
  'TCHCARE_AMT', 'TMDPAY',

  # Assets (individual-level; shared assets divided equally)
  'TVAL_AST', 'TVAL_BANK', 'TVAL_BOND', 'TVAL_BUS', 'TVAL_ESAV',
  'TVAL_HOME', 'TVAL_RENT', 'TVAL_RE', 'TVAL_RET',
  'TVAL_STMF', 'TVAL_VEH', 'TVAL_OTH',

  # Program participation
  'RFSCOV', 'RSNAP_MNYN', 'RTANF_MNYN', 'RSSI_MNYN', 'RWICYN',

  # ABAWD work requirement variables
  'TMWKHRS', 'EWELACTV2_1', 'EWELACTV2_2', 'EWELACTV2_3', 'EWELACTV4'
)

patterns <- c(
  "^RREL", "^ET2REL", "^SNAP", "^TEHC",
  "^EOWN", "^EJSOWN", "^EJOOWN", "^EOOWN",
  "^TJS", "^TJO", "^VAL", "^EVEH", "^TVEH"
)

pattern_vars <- unique(unlist(lapply(patterns, function(p) grep(p, all_cols, value = TRUE))))
keep_vars    <- unique(c(exact_vars, pattern_vars))
keep_vars    <- keep_vars[keep_vars %in% all_cols]

pu <- fread(ds, sep = "|", select = keep_vars)
names(pu) <- toupper(names(pu))

dw <- "data/rw2019.csv"
rw <- fread(dw, sep = "|")
names(rw) <- toupper(names(rw))

sipp2019 <- left_join(pu, rw, by = c("SSUID", "PNUM", "MONTHCODE"))
```

## Filter to July 2019 and generate

```{r}
sipp_0719 <- sipp2019 |>
  filter(RIN_UNIV == 1 & MONTHCODE == 7) |>
  mutate(hhid = paste0(as.character(SSUID), as.character(SHHADID))) |>
  group_by(hhid) |>
  mutate(snap_split = ERELRPE %in% 16:18, # splitting cases that are roommates/boarders/other non-family into separate SNAP unit
         split_id = if_else(snap_split, cumsum(snap_split), NA_integer_),
         snapuid = case_when(!snap_split ~ 1,
         snap_split ~ split_id + 1)
) |> ungroup() |>
  mutate(snap_hhid = paste0(hhid, snapuid))

# Brief checks to make sure split went correctly
sipp_0719 |>
  group_by(hhid) |>
  summarise(n_snap_units = n_distinct(snap_hhid)) |>
  count(n_snap_units)

sipp_0719 |>
  group_by(hhid) |>
  filter(n_distinct(snap_hhid) > 1) |>
  arrange(hhid, snap_hhid) |>
  select(hhid, ERELRPE, snap_hhid) |>
  ungroup() |>
  slice_head(n = 50)

```

## FPL and Deduction Limits

```{r}
# 2019 monthly FPL (contiguous 48 states and DC)
# https://aspe.hhs.gov/topics/poverty-economic-mobility/poverty-guidelines/prior-hhs-poverty-guidelines-federal-register-references/2019-poverty-guidelines 
fpl_monthly    <- c(1041, 1410, 1778, 2146, 2515, 2883, 3251, 3620)
fpl_additional <- 369

# 2019 standard deductions by household size (caps at 6)
# https://cvlas.org/wp-content/uploads/2022/08/supp-nutrition-assist-prog-SNAP-2019.pdf
std_ded_tbl <- c(167, 167, 167, 178, 208, 238)

# Excess shelter deduction cap for non-elderly/disabled households
shelter_cap <- 552

get_fpl <- function(n){
  fpl_monthly[pmin(n, 8)] + pmax(n - 8, 0) * fpl_additional
}
get_std_ded <- function(n){
  std_ded_tbl[pmin(n, 6)]
}
```

## Aggregate to Household-Level SNAP Filing Units

SNAP filing units are people who live and purchase food together. Given
that SIPP doesn't account for people who consume food together, we use
SSUID to account for the pseudo-economic unit.

[It's important to note the SIPP is very well known for undercounting
income
information.](https://www.aeaweb.org/articles?id=10.1257/jep.29.4.199)

```{r}
# Aggregation
hh_sipp <- sipp_0719 |>
  group_by(snap_hhid) |>
  summarize(
    hh_size = n(),
    has_elderly = any(TAGE >= 60, na.rm = TRUE),
    has_disabled = any(RDIS == 1, na.rm = TRUE),
    has_ssi = any(RSSI_MNYN == 1, na.rm = TRUE),
    has_tanf = any(RTANF_MNYN == 1, na.rm = TRUE),
    # Income
    hh_gross_inc = sum(TPTOTINC, na.rm = TRUE) * 1.25, # sipp undercounts apply this factor
    hh_earn = sum(TPEARN, na.rm = TRUE) * 1.25, # sipp undercounts apply this factor
    # Deductions
    child_sup_ded = sum(TAMOUNTPAID, na.rm = TRUE),
    child_care_ded = sum(TPAYWK, na.rm = TRUE) * 4,    
    medical_costs = sum(TMDPAY, na.rm = TRUE),
    # Shelter (December of reference year as proxy)
    shelter_costs = coalesce(first(na.omit(THVAL_RMU)), 0),
    # Countable assets
    countable_assets = sum(TVAL_BANK, na.rm = TRUE) + sum(TVAL_BOND, na.rm = TRUE) +
                       sum(TVAL_BUS,  na.rm = TRUE) + sum(TVAL_STMF, na.rm = TRUE) +
                       sum(TVAL_RENT, na.rm = TRUE) + sum(TVAL_RE,   na.rm = TRUE) +
                       sum(TVAL_OTH,  na.rm = TRUE),
    vehicle_assets = sum(TVAL_VEH, na.rm = TRUE),
    tveh1_val = first(TVEH1VAL),
    tveh2_val = first(TVEH2VAL),
    # ABAWD flags
    has_children = any(TAGE < 18, na.rm = TRUE),      
    youngest_child_age = if (any(TAGE < 26, na.rm = TRUE)) 
                           as.numeric(min(TAGE[TAGE < 26], na.rm = TRUE))
                         else NA_real_,
    has_abawd_age = any(TAGE >= 18 & TAGE <= 49 & # baseline 18-49 flag
                       (is.na(RDIS) | RDIS != 1), na.rm = TRUE),
    has_adult_nondis_50_59 = any(TAGE >= 50 & TAGE <= 59 &
                                (is.na(RDIS) | RDIS != 1), na.rm = TRUE),
    has_adult_nondis_60_65 = any(TAGE >= 60 & TAGE <= 65 &
                                (is.na(RDIS) | RDIS != 1), na.rm = TRUE),
    max_work_hrs = max(c(0, TMWKHRS[TAGE >= 18 &         
                       (is.na(RDIS) | RDIS != 1)]), na.rm = TRUE),
    has_any_citizen = any(ECITIZEN %in% c(1, 2), na.rm = TRUE), 
    state_fips = first(TEHC_ST),
    weight = first(WPFINWGT),
    snap_actual = any(RSNAP_MNYN == 1, na.rm = TRUE)
  )
```

## Eligibility Function

State-level BBCE and vehicle rules are looked up internally from
`snapdb` using `state_fips`. We hardcoded three national or other
baseline level policy levers which can then be modified for simulation.

They are income, asset, and abawd work policy levers.

```{r}
snap_elig_sim <- function(df,
    # Note all of these can be changed as simulations: Income 
    gross_pct = 1.30,     # gross income limit as fraction of FPL (federal default: 1.30)
    net_pct = 1.00,     # net income limit as fraction of FPL   (federal default: 1.00)
    use_bbce= TRUE,     # apply state BBCE income extensions
    # Asset 
    elim_asset = FALSE,  # TRUE = no asset test for anyone
    asset_limit = NULL,   # set a uniform asset limit in dollars; NULL = federal/BBCE default
    # Work 
    work_req = NULL, # NULL = use state rules; "full_exemption"; "no_exemption"
    abawd_max_age = 49,   # upper age of ABAWD range (federal default: 49)
    min_work_hrs = 20,   # minimum weekly hours to meet work requirement 
    dep_age = 18    # child must be under this age to exempt adult from ABAWD 
) { with(df, 
         {
  # State snap policy
  sp <- snapdb[match(state_fips, snapdb$state_fips), ]

  # FPL thresholds
  fpl       <- get_fpl(hh_size)
  gross_lmt <- fpl * gross_pct
  net_lmt   <- fpl * net_pct
  fed_asset_lmt <- if_else(has_elderly | has_disabled, 4500, 3000)
  if (!is.null(asset_limit)) fed_asset_lmt <- rep(asset_limit, length(hh_size))

  # Net income calculation
  std_ded <- get_std_ded(hh_size)
  earn_ded <- hh_earn * 0.20
  net_pre_shelter <- pmax(hh_gross_inc - std_ded - earn_ded - child_sup_ded - child_care_ded, 0)
  shelter_excess  <- pmax(shelter_costs - 0.50 * net_pre_shelter, 0)
  shelter_ded <- if_else(has_elderly | has_disabled,
                             shelter_excess, pmin(shelter_excess, shelter_cap))
  medical_ded <- if_else(has_elderly | has_disabled, pmax(medical_costs - 35, 0), 0)
  net_inc <- pmax(net_pre_shelter - shelter_ded - medical_ded, 0)

  # Categorical eligibility (SSI/TANF bypasses income and asset tests)
  cat_elig <- has_ssi | has_tanf

  # BBCE — look up from state policy; disabled when use_bbce = FALSE
  bbce_on <- use_bbce & !is.na(sp$bbce) & sp$bbce == 1
  bbce_applies <- bbce_on & !(coalesce(sp$bbce_child, 0L) == 1L & !has_children)

  gross_lmt_bbce <- if_else(bbce_applies & coalesce(sp$bbce_inclmt, -9L) > 0,
                             fpl * (sp$bbce_inclmt / 100), gross_lmt)
  gross_lmt_eld <- if_else(bbce_applies & coalesce(sp$bbce_elddisinclmt, -9L) > 0,
                             fpl * (sp$bbce_elddisinclmt / 100), gross_lmt_bbce)
  gross_lmt_eff <- if_else(has_elderly | has_disabled, gross_lmt_eld, gross_lmt_bbce)

  asset_test_on <- !(bbce_applies & !is.na(sp$bbce_asset) & sp$bbce_asset == 1)
  asset_lmt_eff <- case_when(
    bbce_applies & coalesce(sp$bbce_a_amt, 0) > 0 ~ sp$bbce_a_amt * 1000,
    TRUE ~ fed_asset_lmt
  )
  if (elim_asset) asset_test_on <- rep(FALSE, length(hh_size))

  # Vehicle countable value — look up state vehicle exclusion rules
  vehicle_countable <- case_when(
    sp$vehexclall == 1 ~ 0,
    sp$vehexclone == 1 ~ pmax(vehicle_assets - coalesce(tveh1_val, 0), 0),
    TRUE ~ vehicle_assets
  )
  net_worth <- countable_assets + vehicle_countable

  # Eligibility tests
  gross_elig <- hh_gross_inc <= gross_lmt_eff
  net_elig   <- net_inc      <= net_lmt
  asset_elig <- !asset_test_on | cat_elig | (net_worth <= asset_lmt_eff)

  eff_work_req <- if (!is.null(work_req)) rep(work_req, length(hh_size)) else sp$work_req
  has_abawd_age_eff  <- has_abawd_age |
                        (abawd_max_age >= 55 & has_adult_nondis_50_59) |
                        (abawd_max_age >= 60 & has_adult_nondis_60_65)
  has_qualifying_dep <- !is.na(youngest_child_age) & youngest_child_age < dep_age
  meets_work_eff <- max_work_hrs >= min_work_hrs
  has_abawd <- has_abawd_age_eff & !has_qualifying_dep & !has_ssi & !has_tanf
  work_elig <- eff_work_req == "full_exemption" | !has_abawd | meets_work_eff

  citizen_elig <- coalesce(has_any_citizen, TRUE) | cat_elig

  (cat_elig | (gross_elig & net_elig)) & asset_elig & work_elig & citizen_elig
})}

```


## Scenario Functions

```{r}
run_scenario <- function(data, scenario_name, ...) {
  data |>
    mutate(snap_eligible = snap_elig_sim(pick(everything()), ...),
           scenario = scenario_name)
}

summarize_scenario <- function(data) {
  data |>
    ungroup() |>
    summarize(
      scenario = first(scenario),
      n_eligible_weighted = sum(snap_eligible * weight, na.rm = TRUE),
      n_total_weighted = sum(weight, na.rm = TRUE),
      pct_eligible= round(weighted.mean(snap_eligible, weight, na.rm = TRUE) * 100, 1))
}
```

## Run Scenarios

Roughly [82% of SNAP eligible
households](https://www.fns.usda.gov/research/snap/state-participation-rates/2019#:~:text=Key%20Findings,Page%20updated:%20April%2018%2C%202025)
participate in snap in 2019. [July 2019 had 19 million
households](https://www.fns.usda.gov/pd/supplemental-nutrition-assistance-program-snap).
As such assuming the SIPP properly captures income, asset information,
then we should have roughly 23.5 million households.

```{r}
# Baseline: 2019 state rules as-is
baseline <- run_scenario(hh_sipp, "Baseline (2019)")

# Income
income_200pct  <- run_scenario(hh_sipp, "Gross limit 200% FPL",
                               gross_pct = 2.00)

income_no_bbce <- run_scenario(hh_sipp, "Federal rules only (no BBCE)",
                               use_bbce = FALSE)

# Assets
asset_elim <- run_scenario(hh_sipp, "Eliminate asset test",
                           elim_asset = TRUE)

asset_10k  <- run_scenario(hh_sipp, "Raise asset limit to $10k",
                           asset_limit = 10000)

# Work requirement scenarios 
work_waive <- run_scenario(hh_sipp, "Waive ABAWD nationally",
                             work_req = "full_exemption")

work_enforce <- run_scenario(hh_sipp, "Enforce ABAWD nationally",
                             work_req = "no_exemption")

work_30hrs <- run_scenario(hh_sipp, "30-hour requirement", min_work_hrs = 30)

work_dep5    <- run_scenario(hh_sipp, "Dep age cutoff 5", dep_age = 5)

work_age65 <- run_scenario(hh_sipp, "ABAWD age 18-65", abawd_max_age = 65)
```

## Results

```{r}
results <- bind_rows(
  summarize_scenario(baseline),
  summarize_scenario(income_200pct),
  summarize_scenario(income_no_bbce),
  summarize_scenario(asset_elim),
  summarize_scenario(asset_10k),
  summarize_scenario(work_waive),
  summarize_scenario(work_enforce),
  summarize_scenario(work_30hrs),
  summarize_scenario(work_dep6),
  summarize_scenario(work_age65),
) |>
  mutate(
    change_from_baseline = n_eligible_weighted - n_eligible_weighted[scenario == "Baseline (2019)"],
    pct_change = round(100 * change_from_baseline / n_eligible_weighted[scenario == "Baseline (2019)"], 1)
  )

results
```
